---
title: "World after Covid: Method & Analyses"
author: "Igor Grossmann"
date: "Jan 11,2021"
output:
  rmdformats::html_clean:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table header
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r,echo=FALSE,include=FALSE}

#'second, load packages
#library(haven)
#BiocManager::install("NetPathMiner") #if need to install for drawing legend better.
#library(NetPathMiner)
library(rmdformats)
library(dplyr)
library(irr) # to get reliability estimates for codes
library(tidyr)
library(ggplot2)
library(ggpubr)
library(ggthemes)
library(ggsci)
library(plotly)
library(reshape2)
library(ggtext)
library(lubridate)
library(stringr)
library(lattice)
library(quanteda)
library(readtext)
library(psych)
library(tm)
library(igraph)
library(ggraph)
library(cluster)
library(factoextra)
library(RColorBrewer)
library(purrr)
library(lme4)
library(emmeans)
library(packcircles)
library(viridis)
library(ggiraph)
library(edgebundleR)
#library(GGally)
library(GPArotation)
library(heatmaply)
library(vcd)
library(tibble)
library(lme4)
library(ggwordcloud)
library(showtext)
library(car)


```

```{r,echo=FALSE,include=FALSE}
setwd("C:/Users/igrossma/Documents/GitHub/wac") #change to your preferred working directory hosting the csv files
codes.data<-read.csv("codes.csv")
transcripts<-read.csv("transcripts.csv")
codes.outsider<-read.csv("outside_dialectic_codes.csv")

all.data<- transcripts %>% full_join(codes.data, by = c("Name"))
all.data<- all.data %>% full_join(codes.outsider, by = c("Name"))

all.data$date<-parse_date_time(all.data$RecordingDate, orders = c("dmy", "mdy",("ymd"))) #recode date as date


all.data<-all.data[order(as.Date(all.data$date, format="%Y/%m/%d")),] #resort by date


all.data$Week <- as.Date(cut(all.data$date,
                        breaks = "week",
                        start.on.monday = T)) # changes weekly break point to Sunday


all.data %>% group_by(date) %>% count() %>% ggplot() + geom_line(aes(date,n, group = 1)) #how many interviews on what days

all.data <- all.data %>% mutate(month = month(date))

all.data <- all.data %>% mutate(week = week(date))


all.data %>% group_by(month) %>% count()  #%>% ggplot() + geom_bar(aes(month,n), stat ='identity')

#how many interviews on different weeks
all.data %>% group_by(Week) %>% count()  #%>% ggplot() + geom_bar(aes(Week,n), stat ='identity')

#'get tokens for each Q.

all.data$Q1.text<-as.character(all.data$Q1)
Q1.corpus <- corpus(all.data, text_field = "Q1.text")
all.data$tokens.Q1<-ntoken(Q1.corpus)

all.data$Q2.text<-as.character(all.data$Q2)
Q2.corpus <- corpus(all.data, text_field = "Q2.text")
all.data$tokens.Q2<-ntoken(Q2.corpus)

all.data$Q3.text<-as.character(all.data$Q3)
Q3.corpus <- corpus(all.data, text_field = "Q3.text")
all.data$tokens.Q3<-ntoken(Q3.corpus)

all.data$Q4.text<-as.character(all.data$Q4)
Q4.corpus <- corpus(all.data, text_field = "Q4.text")
all.data$tokens.Q4<-ntoken(Q4.corpus)

all.data$Q5.text<-as.character(all.data$Q5)
Q5.corpus <- corpus(all.data, text_field = "Q5.text")
all.data$tokens.Q5<-ntoken(Q5.corpus)

```


# The method

## Recruitment

Since June 2020, I have interviewed 57 of the [top social, behavioral, and political scientists](https://github.com/grossmania/wac/blob/main/participants.md), along with several eminent philosophers, wisdom scientists, and futurists. First, the advisory board and I nominated 153 top scholars from around the world. From the list of contacted nominees, 65% responses to the initial inquiry to participate in the interview series. Among those who responded, 37% declined. I ended up scheduling 38% of the scholars on the initial list -- a reasonable response rate considering the pandemic-related challenges and the nature of the project.\

## Quantifying narratives

To quantify themes in interviews, we relied on a novel **Multi-step Bottom-up/Top-down Cross-validation** method for analyzing and cross-referencing unstructured qualitative data. in the method, we first rely on classic *grounded approach*, with two scholars going through the first set of 30 interviews June-July) to identify unique themes for each question in an iterative fashion. Hereby, we followed the following guidelines:\

*   Each theme should be present at least twice across interviews;\
*   Themes should have minimum overlap. In other words, they should not be directly reducible, though they could show natural dependencies (e.g., "importance of social connections" and "social support").

In this initial phase, two independent raters, one of whom was blind to the identity of the interviewees, coded statements on prevalence of determined themes. Initial reliability was good (over 90% agreement), with disagreements resolved via group discussion with the senior scholar. Following iterative procedure, in case additional themes were identified that were not covered by the original categories, they were added to the codebook and statements were recoded for presence of the category.\
\
In the second stage, after interviews were extended into September - December 2020, another two coders (one of whom coded the initial set of 30 interviews) coded the new batch of interviews. Once again, agreement was high (90+%), with disagreements resolved in a group discussion. In this stage, coders identified several additional themes, which were again added to the codebook, with all interviewed cross-examined for presence of these themes. The procedure was repeated four times, until we identified the final list of themes for each question. Here, we also included some simple-occurrence themes if they were fully distinct and addressed the questions.\

Coding open-ended interviews is inherently subjective. Even in the presence of high reliability between coders (as in our case), validity of the coding may be compromised due to various additional factors (e.g., a particular sentiment in a response, agreement with the opinion raised in the interview response). To address this issue, we introduce a novel *top-down cross-validation approach*: \

*   A new, unbiased person blind to identity of interviewees reviews codes and respective transcripts, with the task to identify one key sentence [or key phrases, in case the theme is not captured by a single sentence] from each person’s response to represent their code, guided by the codebook definitions;
*   Two further individuals, incl. the senior author of the project, review these key sentences and flag any categories that required adjustment.\ 

The idea behind this cross-validation approach is that this top-down, bird’s eye view allows for greater clarity when matching codes and themes compared to the classical grounded analysis. By matching each code to a core statement/phrase(s), one introduces extra rigor when evaluating each and every code. Indeed, in the process of such cross-validation, several minor inconsistencies were spotted and corrected prior to conducing subsequent analyses.

## Quantifying reasoning style

Prior research on forecasting suggests that a certain cognitive style may be more conducive for accurate forecasting of [geopolitical events](https://doi.org/10.1177/1745691615577794), [affect](https://doi.org/10.1037/0022-3514.78.5.821) and emotions toward close others in [social conflict situations](https://psyarxiv.com/a5fgu). Specifically, research suggests that superior forecasters tend to show greater likelihood of embracing:\

*   an **outsider viewpoint** and consider additional baserate information (rather than focus on the focal event alone);\
*   more complex, **dialectical thinking** (aspects of which are central to the notions of [integrative complexity](https://doi.org/10.1037/0022-3514.49.6.1565) and [wisdom](https://doi.org/10.1080/1047840X.2020.1750917)) -- i.e., recognize the uncertainty and qualify forecasts by expressing multi-determined nature of predictions and considering both positive and negative aspects of the same forecast.\

I sought to quantify the extent to which top experts in behavioral and social sciences apply these aspects of reasoning in their reflections. To this end, my team focused on the forecast-related questions (Q1: positive consequences / Q3: negative consequences). Here, we could categorize responses as those invoking *outsider viewpoint* and *dialectical reasoning* and compare % frequencies and type of forecasts among these two groups. A codebook for both categorization types is [here](https://github.com/grossmania/wac/blob/main/codebook_outsider.md). Because dialectical reasoning categorization involved both questions, each participant received only one code (yes=1 / no = 0). For outsider viewpoint, we separately categorized positive and negative responses, which allowed us to compare likelihood of invokingoutsider-view information across questions. Two research assistants, who did not engage in coding prior categories), independently categorized responses on both categories. As analyses below show, inter-rater reliability was medium-large (rules of thumb for Cohen's kappa suggest *h* =.5 as medium effect size and *h* = .8 as large effect size). Disagreements were resolved in a discussion with the senior author and another co-author on the project.

```{r reliability of codes,warning=FALSE,message=FALSE}
kappa2(all.data[c("Q1_Outside_CC","Q1_Outside_OM")]) #Question 1: Outsider Viewpoint
kappa2(all.data[c("Q3_Outside_CC","Q3_Outside_OM")]) #Question 3: Outsider Viewpoint
kappa2(all.data[c("Dialectic_CC","Dialectic_OM")]) #Dialectical Thinking

```

# Descriptives

Beyond US, scientists from Australia, Canada, Korea, Japan, HK, Russia, Germany, Spain, Switzerland, Israel, UK. \
[The list of luminaries](https://github.com/grossmania/wac/blob/main/participants.md) includes past and present presidents/chairs of APS, APA, Psychonomic Society, Society for Personality and Social Psychology, International Association for Cross-cultural Psychology, Society for Affective Science, Cultural Evolution Society, Human Behavior and Evolution Society, along with numerous members of the US National Academy of Sciences, US NAtional Academy of Education, US National Academy of Engineering, US National Academy of Medicine, Royal Canadian Academy, German National Academy of Sciences Leopoldina,  Academy of the Social Sciences in Australia, and American Academy of Arts and Sciences.

## Breakdown by gender, location, and expertise 

```{r,echo=FALSE,warning=FALSE,message=FALSE}
#gender breakdown
all.data %>% group_by(Gender)%>% 
  summarise(count = n() ) %>%
  mutate( prop = count / sum(count)*100 )%>% ggplot() + geom_bar(aes(Gender,prop, fill=Gender), stat ='identity')+scale_fill_d3()+theme(legend.position = "none")+labs(y="%", x="")+theme_pubclean()+theme(legend.position="none")


#location breakdown
all.data %>% group_by(Location)%>% 
  summarise(count = n() ) %>%
  mutate( prop = count / sum(count)*100 )%>% ggplot() + geom_bar(aes(Location,prop, fill=Location), stat ='identity')+scale_fill_d3()+theme(legend.position = "none")+labs(y="%", x="Location")+theme_pubclean()+theme(legend.position="none")

```

*SC - socio-cultural expertise*\
*Non-SC - no socio-cultural expertise*

Expertise is roughly equally distributed.



```{r,echo=FALSE,warning=FALSE,message=FALSE}
#expertise breakdown
all.data %>% group_by(Field)%>% 
  summarise(count = n() ) %>%
  mutate( prop = count / sum(count)*100 )%>% ggplot() + geom_bar(aes(Field,prop, fill=Field), stat ='identity')+scale_fill_d3()+theme(legend.position = "none")+labs(y="%", x="Expertise (Socio-cultural vs. Rest)")+theme_pubclean()+theme(legend.position="none")


#wordcloud of subfields
table(all.data$Domain)
all.data$Domain_B[all.data$Domain_B==""]<-NA
table(all.data$Domain_B)
domain<-as.data.frame(table(all.data$Domain))
domain2<-as.data.frame(table(all.data$Domain_B))
domains<-full_join(domain,domain2,by = c("Var1"))
domains$Freq<-rowSums(domains[c("Freq.x","Freq.y")], na.rm=T)
domains$Type<-c(1,2,3,4,5,2,6,7,8,6,8,6,2,3,4,6,3,6,9,10,3,7,8,6,6,2,6,4,2,6,3,4)
ggplot(domains, aes(label = Var1, size = Freq, color=as.factor(Type))) +
  geom_text_wordcloud() +
  scale_size_area(max_size = 15) +scale_color_d3(palette="category10")+
  theme_minimal()

```

Only 20% were familiar with wisdom research (either in philosophy or in psychology).


# Cognitive Style Differences among Experts: Breakdown by dialectical thinking & outsider viewpoint

Almost half of the experts invoked dialectical thinking when reflecting on future consequences of the pandemic. In other words, they pointed out that the same theme could have both positive and negative consequences. This way, they also communicated their uncertainty in the downstream impact of selected themes.\

Less than a third of experts invoked baserates or other outsider view information when qualifying their prediction for positive (25%) or negative forecasts (27%). Ten percent of experts brought baserate or other outsider view considerations in reflection on both scenarios. These was no substantial difference between consideration of baserates when forecasting positive vs. negative consequences.


```{r,echo=FALSE,warning=FALSE,message=FALSE}
#dialectical thinking

all.data$Dialectic_Final<-as.factor(all.data$Dialectic_Final)

all.data %>% group_by(Dialectic_Final)%>% 
  summarise(count = n() ) %>%
  mutate( prop = count / sum(count)*100 )%>% ggplot() + geom_bar(aes(Dialectic_Final,prop, fill=Dialectic_Final), stat ='identity')+scale_fill_d3()+theme_pubclean()+scale_x_discrete(labels=c("No","Yes"))+theme(legend.position = "none")+labs(y="%", x="Dialectical Thinking")

#outsider viewpoint
all.data$'Positive Consequences'<-as.factor(all.data$Q1_Outside_Final)
all.data$'Negative Consequences'<-as.factor(all.data$Q3_Outside_Final)

all.data %>% filter(!is.na(Q3_Outside_Final))%>% 
  pivot_longer(c("Positive Consequences","Negative Consequences"),names_to = "Outside.type", values_to = "Code")%>% 
  ggplot(aes(Code, group = Outside.type)) + 
          geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
          scale_y_continuous(labels=scales::percent) +
          facet_grid(~Outside.type)+scale_x_discrete(labels=c("No","Yes"))+scale_fill_d3()+theme_pubclean()+theme(legend.position = "none")+labs(y="Relative Rrequencies", x="Outsider Viewpoint")

prop.table(table(all.data$Q1_Outside_Final))
prop.table(table(all.data$Q3_Outside_Final))


#outsider viewpoint in total (pos and neg consequences combined)

all.data$Outsider_total<-as.factor(rowSums(all.data[c("Q1_Outside_Final","Q3_Outside_Final")],na.rm=T))

prop.table(table(all.data$Outsider_total))

all.data %>% group_by(Outsider_total)%>% 
  summarise(count = n() ) %>%
  mutate( prop = count / sum(count)*100 )%>% ggplot() + geom_bar(aes(Outsider_total,prop, fill=Outsider_total), stat ='identity')+scale_fill_d3()+theme(legend.position = "none")+labs(y="%", x="Outsider Viewpoint in Total")+theme_pubclean()+theme(legend.position="none")


```
About 22 % of participants wrote on the topic of wisdom whereas the majority did not know much about wisdom scholarship, with some scholars expolicitly stating they don't know what wisdom refers to, in which case they were referred to the broad definition concerning "attitudes, strategies, and behaviors that can help master the pandemic."


```{r,echo=FALSE,warning=FALSE,message=FALSE}
#expertise breakdown
all.data %>% group_by(FamiliarWisdom)%>% 
  summarise(count = n() ) %>%
  mutate( prop = count / sum(count)*100 )%>% ggplot() + geom_bar(aes(FamiliarWisdom,prop, fill=FamiliarWisdom), stat ='identity')+scale_fill_d3()+theme(legend.position = "none")+labs(y="%", x="Familiarity with Wisdom Research?")+theme_pubclean()+theme(legend.position="none")


```

# Positive Consequences

## Summary

### A great degree of variability in themes

We identified 20 distinct themes.\
Most themes mentioned by less than 10% of the interviewees.\
Only three themes were mentioned by at least ten people: greater (appreciation of) social connectedness,  opportunity for political engagement/structural change viewed, and solidarity.\

* Greater (appreciation of) social connectedness, and opportunity for political engagement/structural change viewed, and solidarity as most pos outcomes. \
* Additionally: opportunity to re-evaluate habits and embracing new technology. \
* Greater (appreciation of) social connectedness and solidarity, as well as resilience more frequent among women, and opportunity for political engagement/structural change and  reconsider habits more frequent among men. Reflecting gender roles?\
* Socio-cultural experts are more likely to mention social connectedness and opportunity for political engagement/structural change, whereas non-SC experts folked more on reconsidering habits.\

## Frequency Chart

```{r,echo=FALSE,warning=FALSE,message=FALSE}
Q1 <- c("shared.humanity_q1","critical.thinking_q1","personal.resilience_q1","gratitude_q1",
        "nature_q1","improved.care.for.elders_q1","embrace.new.tech_q1","prosocial.behavior_q1",
        "social.connectedness_q1","improved.work.life.balance_q1","increased.interest.in.science_q1",
        "learning.from.pandemics_q1","political.engagement...structural.change_q1",
        "optimism...positivity_q1","reconsidering.habits_q1",
        "living.in.the.moment_q1","health...well.being_q1","solidarity_q1","bipartisanship.and.international.cooperation_q1","sympathy...compassion_q1") 


Q1.q <- c("Shared Humanity","Critical Thinking","Resilience","Gratitude",
          "Nature","Care for Elders","Embrace New Tech","Prosocial Behavior",
          "Social Connectedness","Work-Life Balance","Science Interest","Learning from Pandemics","Polit./Structure-Change",
          "Optimism/Positivity","Reconsider Habits","Live\nin the Moment","Health/Wellbeing","Solidarity","Political Cooperation","Sympathy&Compassion") 

Q1.type <- c("Meta-cognition","Meta-cognition","Psych.\nWellbeing","Psych.\nWellbeing",
             "Nature","Prosocial","Work","Prosocial","Prosocial","Work",
             "Quest\nfor Knowledge","Quest\nfor Knowledge","Societal",
             "Psych.\nWellbeing","Psych.\nWellbeing","Psych.\nWellbeing",
             "Psych.\nWellbeing","Prosocial","Societal","Meta-cognition") 

Q1.data<-all.data %>% 
  summarise(across(shared.humanity_q1:sympathy...compassion_q1, mean))



colnames(Q1.data)[colnames(Q1.data)      # Rename two variable names
                     %in% Q1] <- Q1.q

Q1.data.long <- Q1.data %>% 
  pivot_longer(everything(),names_to = "Theme", values_to = "Proportion")
Q1.summ<-cbind(Q1.data.long, Q1.type)
colnames(Q1.summ) <- c("Theme","Proportion","Type")  # change column names
Q1.summ <- Q1.summ[order(Q1.summ$Proportion), ]  # sort
Q1.summ$Theme <- factor(Q1.summ$Theme, levels = Q1.summ$Theme)  # to retain the order in plot
Q1.summ$Percent<-Q1.summ$Proportion*100
Q1.plot<-ggplot(Q1.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() +
  geom_bar(stat="identity", width=.5) + ggtitle("Positive Consequences")+theme_pubclean()+scale_fill_d3()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="%") 
Q1.plot
```

## How many themes per person?

Did people just report 1-2 themes or a great number of themes? How do such trends vary across people?
```{r,echo=FALSE,warning=FALSE,message=FALSE}
all.data$Q1.pp<-rowSums(all.data[Q1])

ggplot(all.data, aes(x = Q1.pp)) +   geom_vline(xintercept=median(all.data$Q1.pp), color="red")+
  geom_bar()+theme_pubclean()+scale_fill_d3()+labs(x="Number of Themes per Person",y="Count")
           
```

The majority of people mention only one or two themes here. The median was two themes (as indicated by vertical red line). One-fifth mentioned three themes, and only 3 people mentioned 4 themes (and one person mentioned five themes). 

## Frequencies by Gender and Field (Socio-cultural vs. Other fields)

What is important here is that the focus on social-structural issues was not only present among experts in social psychology or cultural matters but also among experts studying other areas of psychology or political science.


```{r,echo=FALSE,warning=FALSE,message=FALSE}
##By gender
Q1.data.G<-all.data %>% group_by(Gender)%>% 
  summarise(across(shared.humanity_q1:sympathy...compassion_q1, mean))

colnames(Q1.data.G)[colnames(Q1.data.G)      # Rename two variable names
                     %in% Q1] <- Q1.q
Q1.data.G.long <- Q1.data.G %>% 
  pivot_longer(-Gender,names_to = "Theme", values_to = "Proportion")
Q1.G.summ<-cbind(Q1.data.G.long, Q1.type)
colnames(Q1.G.summ) <- c("Gender","Theme","Proportion","Type")  # change column names
Q1.G.summ$Percent<-Q1.G.summ$Proportion*100
Q1.G.plot<-ggplot(Q1.G.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Gender)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Positive Consequences")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="%")
Q1.G.plot

##By socio-cultural vs. non-socio-cultural expertise
Q1.data.SC<-all.data %>% group_by(Field)%>% 
  summarise(across(shared.humanity_q1:sympathy...compassion_q1, mean))

colnames(Q1.data.SC)[colnames(Q1.data.SC)      # Rename two variable names
                     %in% Q1] <- Q1.q

Q1.data.SC.long <- Q1.data.SC %>% 
  pivot_longer(-Field,names_to = "Theme", values_to = "Proportion")
Q1.SC.summ<-cbind(Q1.data.SC.long, Q1.type)
colnames(Q1.SC.summ) <- c("Field","Theme","Proportion","Type")  # change column names
Q1.SC.summ$Percent<-Q1.SC.summ$Proportion*100
Q1.SC.plot<-ggplot(Q1.SC.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Field)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Positive Consequences")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="%")
Q1.SC.plot
```

## Frequencies by Cognitive Style: Dialectical Thinking and Consideration of Outsider Viewpoint

It appears that **people invoking dialecticism** in their reasoning were similar in their positive forecasts as those who did not show dialecticism.\
The only noticeable exception is the much greater emphasis on social connectedness among scholars showing dialecticism in their reasoning.\
\
It also appears that **people invoking baserate/outsider view information** in their reasoning were also quite similar in their positive forecasts as those who did not mention baserate info. There are some differences, but they may be due to differences in sub-sample size.\
The only noticeable exceptions are:\

 - greater focus on *resilience* and *wellbeing* among *baseraters* and greater focus on *reconsidering habits* among *non-baseraters*.\ 

```{r,echo=FALSE,warning=FALSE,message=FALSE}
#Dialectical thinking

Q1.data.Dia<-all.data %>% group_by(Dialectic_Final)%>% 
  summarise(across(shared.humanity_q1:sympathy...compassion_q1, mean))

colnames(Q1.data.Dia)[colnames(Q1.data.Dia)      # Rename two variable names
                     %in% Q1] <- Q1.q
Q1.data.Dia$Dialecticism[Q1.data.Dia$Dialectic_Final==0]<-"Not Dialectical"
Q1.data.Dia$Dialecticism[Q1.data.Dia$Dialectic_Final==1]<-"Dialectical"

Q1.data.Dia.long <- Q1.data.Dia %>% 
  pivot_longer(-c("Dialectic_Final","Dialecticism"),names_to = "Theme", values_to = "Proportion")
Q1.Dia.summ<-cbind(Q1.data.Dia.long, Q1.type)
colnames(Q1.Dia.summ) <- c("Dialectical Code","Dialecticism","Theme","Proportion","Type")  # change column names
Q1.Dia.summ$Percent<-Q1.Dia.summ$Proportion*100
Q1.Dia.plot<-ggplot(Q1.Dia.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Dialecticism)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Positive Consequences")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="%")
Q1.Dia.plot

Q1.sum.Dia<-all.data %>% group_by(Dialectic_Final)%>% 
  summarise(across(shared.humanity_q1:sympathy...compassion_q1, sum))

#test significance of the effect - prevalence of themes in the dialectical vs. non-dialectical samples.
Q1.data.Dia.long<-all.data %>% pivot_longer(cols=shared.humanity_q1:sympathy...compassion_q1, names_to="Theme",values_to="Presence")

Q1.dia.result<-glmer(Presence ~ Theme*Dialectic_Final + (1 | Name), data = Q1.data.Dia.long, family = binomial)
summary(Q1.dia.result)
Anova(Q1.dia.result,type="III") #no sig difference


#outsider Viewpoint

Q1.data.Outsider<-all.data %>% group_by(Q1_Outside_Final)%>% 
  summarise(across(shared.humanity_q1:sympathy...compassion_q1, mean))

colnames(Q1.data.Outsider)[colnames(Q1.data.Outsider)      # Rename two variable names
                     %in% Q1] <- Q1.q
Q1.data.Outsider$OutsiderV[Q1.data.Outsider$Q1_Outside_Final==0]<-"No Baserate Info (25%)"
Q1.data.Outsider$OutsiderV[Q1.data.Outsider$Q1_Outside_Final==1]<-"Mentioned Baserates (75%)"

Q1.data.Outsider.long <- Q1.data.Outsider %>% 
  pivot_longer(-c("Q1_Outside_Final","OutsiderV"),names_to = "Theme", values_to = "Proportion")
Q1.Outsider.summ<-cbind(Q1.data.Outsider.long, Q1.type)
colnames(Q1.Outsider.summ) <- c("Outsider Code","Baserates","Theme","Proportion","Type")  # change column names
Q1.Outsider.summ$Percent<-Q1.Outsider.summ$Proportion*100
Q1.Outsider.plot<-ggplot(Q1.Outsider.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Baserates)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Positive Consequences")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="%")
Q1.Outsider.plot


```

## PCA

We can address the question of diversity by examining the degree to which scores across themes are reducible to common components. To this end, we perform a principal component analysis (PCA), which works on dichotomous data.\

It appears that 8 component is good start to account for 20 themes for Question 1.\
Keep in mind that reducing 20 items to 8 components is just a 40% reduction and that's not much. Moreover, when we reduce the items to 8 components, the first component explains only 9 % of the variance. Given that each theme by default explains 5% of the variance (1/20), this is not much. \
Furthermore, each of these components is largely based on 1-2 item (loadings > .6), with two items  for TC4 - solidarity & prosocial behavior,  TC6 - shared humanity &bipartisanship, TC3 - critical thinking & increased interest in science, TC5 - personal resilience & optimism/positivity, 
TC2 - health/wellbeing & worklife balance.

```{r,echo=FALSE,warning=FALSE,message=FALSE}

vss(all.data[Q1], rotate="oblimin", fm="pc")

principal(all.data[Q1], nfactors = 8, rotate="oblimin")


```

In short, principal components (note: identical results with minres factor analyses with tetriconic correlation) show substantial diversity, with the top factors shared humanity and critical thinking not accounting for more than 9% of the variance and showing only weak intercorrelation among themselves. And even with 7 components, we don't account for 90% of the variance.\


Three large clusters emerge, similar results are also evident in the dendogram below  .\


## Network Graph

```{r,echo=FALSE,warning=FALSE,message=FALSE}

#first, get correlation matrix., we use spearman and we also control for tokens (length of response for this question)
Q1.cor<-partial.r(all.data,Q1,cs(tokens.Q1),method="spearman")

##################
row.names(Q1.cor) <-Q1.q
colnames(Q1.cor) <-Q1.q

# Keep only meaningful correlations (<.17)
Q1.cor[Q1.cor<0.17] <- 0

# Make an Igraph object from this matrix:
Q1.network <- graph_from_adjacency_matrix( Q1.cor,weighted=T,  mode="undirected", diag=F)


#color palette
coul <- brewer.pal(nlevels(as.factor(Q1.type)), "Dark2")

# Map the color to cylinders
my_color <- coul[as.numeric(as.factor(Q1.type))]

# plot with edges

Q1.net<-plot(Q1.network,
     edge.color=rep(c("red","pink"),5),           # Edge color
     edge.width=seq(1,10),                        # Edge width, defaults to 1
     edge.arrow.size=1,                           # Arrow size, defaults to 1
     edge.arrow.width=1,                          # Arrow width, defaults to 1
     edge.lty=c("solid"),
     edge.label.family = "Times",
     vertex.color =my_color,
     vertex.label.color="black"# Line type, could be 0 or "blank", 1 or "solid", 2 or "dashed", 3 or "dotted", 4 or "dotdash", 5 or "longdash", 6 or "twodash"
     #edge.curved=c(rep(0,5), rep(1,5))            # Edge curvature, range 0-1 (FALSE sets it to 0, TRUE to 0.5)
)

Q1.ceb <- cluster_edge_betweenness(Q1.network) 

dendPlot(Q1.ceb, mode="hclust")
plot(Q1.ceb, Q1.network, axes=T, vertex.label.color="black", edge.width=seq(1,10)) 


```


As we can see above, even after removing negligible correlations (r < .17), cluster analyses on top of the network graphs show five clusters - psych. well-being, civic (poli.structural change, care for elders), solidarity, "pandemic-experience" strategies (social connectedness, change in habits, critical thinking), and adaptive mindsets (shared humanity, focus on nature, optimism/positivity, resilience). Again, this suggests a diversity of topics. Despite forming meaningful dependencies, themes for positive consequences of the pandemic appear largely distinct, except for link between social connectedness and gratitude, linking two networks.

```{r,echo=FALSE,warning=FALSE,message=FALSE}
edgebundle(Q1.network, padding = 120, fontsize = 10)
```

## Convergence vs. divergence of themes over time

I started to interview people in late June, 2020, in the weeks following the BLM protests and riots following George Floyd's death. Over the course of the summer and the fall, numerous other events occurred, including relaxing and re-introducing lockdown policies and regulations in different parts of the world, start of in-person schooling in the fall, the US Presidential election, and its aftermath.\
One way to address the diversity and uncertainty of scientific perspectives on World after Covid and wisdom needed is to examine how consistently themes were mentioned over time. In what follows I do this separately for each question.\

First, I bin frequencies of themes by month. Because only two people provided their interviews in early December, 2020, these individuals will be binned with November scores, such that the scores range from 17 (June), and 19 (July), to six (November).



```{r,echo=FALSE,warning=FALSE,message=FALSE}

all.data$date<-parse_date_time(all.data$RecordingDate, orders = c("dmy", "mdy",("ymd"))) #recode date as date
all.data$month <- as.Date(cut(all.data$date,
                        breaks = "month",
                        start.on.monday = FALSE)) # changes weekly break point to Sunday

all.data <- all.data %>% mutate(month = month(date))

all.data %>% group_by(month) %>% count()  #%>% ggplot() + geom_bar(aes(month,n), stat ='identity')

all.data %>% group_by(date) %>% count()


#bin dates into June, July, Sept/early Oct [pre-lockdown], Oct-Dec 3 [2nd Lock-down/ US Election]
all.data$M<-"4. late Oct+"
all.data$M[all.data$month==6]<-"1. June"
all.data$M[all.data$month==7]<-"2. July"
all.data$M[all.data$month==9]<-"3. Sept/early Oct"
all.data$M[all.data$date=="2020-10-06 UTC"]<-"3. Sept/early Oct"
all.data$M[all.data$date=="2020-10-07 UTC"]<-"3. Sept/early Oct"
all.data$M[all.data$date=="2020-10-09 UTC"]<-"3. Sept/early Oct"

all.data$M.n<-4
all.data$M.n[all.data$month==6]<-1
all.data$M.n[all.data$month==7]<-2
all.data$M.n[all.data$month==9]<-3
all.data$M.n[all.data$date=="2020-10-06 UTC"]<-3
all.data$M.n[all.data$date=="2020-10-07 UTC"]<-3
all.data$M.n[all.data$date=="2020-10-09 UTC"]<-3


table(all.data$M)

Q1 <- c("shared.humanity_q1","critical.thinking_q1","personal.resilience_q1","gratitude_q1",
        "nature_q1","improved.care.for.elders_q1","embrace.new.tech_q1","prosocial.behavior_q1",
        "social.connectedness_q1","improved.work.life.balance_q1","increased.interest.in.science_q1",
        "learning.from.pandemics_q1","political.engagement...structural.change_q1",
        "optimism...positivity_q1","reconsidering.habits_q1",
        "living.in.the.moment_q1","health...well.being_q1","solidarity_q1","bipartisanship.and.international.cooperation_q1","sympathy...compassion_q1") 


Q1.q <- c("Shared Humanity","Critical Thinking","Resilience","Gratitude",
          "Nature","Care for Elders","Embrace New Tech","Prosocial Behavior",
          "Social Connectedness","Work-Life Balance","Science Interest","Learning from Pandemics","Polit./Structure-Change",
          "Optimism/Positivity","Reconsider Habits","Live\nin the Moment","Health/Wellbeing","Solidarity","Political Cooperation","Sympathy&Compassion") 

Q1.type <- c("Meta-cognition","Meta-cognition","Psych.\nWellbeing","Psych.\nWellbeing",
             "Nature","Prosocial","Work","Prosocial","Prosocial","Work",
             "Quest\nfor Knowledge","Quest\nfor Knowledge","Societal",
             "Psych.\nWellbeing","Psych.\nWellbeing","Psych.\nWellbeing",
             "Psych.\nWellbeing","Prosocial","Societal","Meta-cognition") 

Q1.data.M<-all.data %>% group_by(M)%>%
  summarise(across(shared.humanity_q1:sympathy...compassion_q1, sum))


colnames(Q1.data.M)[colnames(Q1.data.M)      # Rename two variable names
                     %in% Q1] <- Q1.q

Q1.data.long.M <- Q1.data.M %>% 
  pivot_longer(-M,names_to = "Theme", values_to = "Count")
Q1.summ.M<-cbind(Q1.data.long.M, Q1.type)
colnames(Q1.summ.M) <- c("Period","Theme","Count","Type")  # change column names
Q1.summ.M <- Q1.summ.M[order(Q1.summ.M$Count), ]  # sort

Q1.data.T<-Q1.summ.M %>% group_by(Period, Type)%>%
    summarise(score=sum(Count))

countByMonth<-all.data %>% group_by(M)%>% 
  summarise(count = n() )

colnames(countByMonth) <- c("Period","N")  # change column names

Q1.data.TT<-full_join(Q1.data.T,countByMonth, by = "Period")
Q1.data.TT$Percent<-(Q1.data.TT$score/Q1.data.TT$N)*100

ggplot(Q1.data.TT, aes(x=Period, y=Percent, group = Type, color=Type, fill=Type)) + 
    geom_line( ) +
    geom_point(shape=21,  size=4) +
   theme_pubclean()+scale_color_d3()+scale_fill_d3()

#BY EACH THEME
#Q1.data.T.ALL<-all.data %>% group_by(M)%>%
#   summarise_at(Q1, funs(sum))
#colnames(Q1.data.T.ALL)[colnames(Q1.data.T.ALL)      # Rename variable names
#                  %in% Q1] <- Q1.q
#
#Q1.data.T.ALL.long <-Q1.data.T.ALL %>%
#    gather(domain, frequency, -c(M), convert = TRUE)
#colnames(Q1.data.T.ALL.long) <- c("Period","Theme","Count")  # change column names
#Q1.data.TT.ALL.long<-full_join(Q1.data.T.ALL.long,countByMonth,, by = "Period")
#Q1.data.TT.ALL.long$Percent<-(Q1.data.TT.ALL.long$Count/Q1.data.TT.ALL.long$N)*100
#ggplot(Q1.data.TT.ALL.long, aes(x=Period, y=Percent, group = Theme, color=Theme, fill=Theme)) + 
#    geom_line( ) +
#  geom_point(shape=21,  size=4) +
#  theme_pubclean()

```

Focus on social issues declines in frequency being mentioned after June/July,  whereas focus on prosociality (social connectedness/solidarity) increases, emerging as dominant in late October right around the time of US election.\
Quest for knowledge (i.e., appreciation of science and willingness to learn from the pandemic) was chiefly mentioned early on, but re-emerged again in late November (after BioNtech vaccine was announced to be possibly 90% effective - Nov 9, 2020).\
Other themes remain largely identical in frequency of being mentioned over time.\

It is also interesting to see how ranking of categories changes over time. Prosocial themes are top at the beginning and esp. in the second wave of the pandemic. Societal issues is close second at the beginning, but move to close to last at the end. \
Psych. Wellbeing emerges as the lead in the summer but becomes relatively less pronounced in late Oct/Nov.\

Though speculative and post-hoc, this observation suggests that for some of the major themes (incl the most frequently mentioned onces), salience of present-day events inform the themes scientists mentioned in their reflections concerning positive consequences -several years- following the pandemic!\


# Negative Consequences

## Summary

### A great degree of variability in themes (again)

We identified 22 distinct themes.\
More than half of themes mentioned by less than 10% of the interviewees.\
Only two themes were mentioned by at least ten people: political conflict and prejudice and racism.\

Generally:\
* Political conflict, prejudice/racism, mistrust, social inequality are among the top negative consequences predicted. \
* Consequences are either social or societal/ with health/well-being being mentioned relatively less. \
* Women more likely to focus on well-being-risks, incl. loneliness and estrangement/alienation, as well as economic hardships. Men more likely to consider prejudice, erosion of democracy. \

## Frequency Chart

```{r,echo=FALSE,warning=FALSE,message=FALSE}
##'QUESTION 3

Q3 <- c("estrangement...alienation_q3","career.disruptions_q3", "decline.in.quality.of.education_q3", "decreased.trust.in.science_q3","decreased.well.being_q3"   ,                     
        "economic.hardship_q3" ,"educational.inequality_q3", "erosion.of.democratic.institutions_q3", "exacerbated.social.inequality_q3","mistrust_q3",                            
        "prejudice...racism_q3", "despair_q3", "prioritize.self.over.others_q3","irrationality_q3","loneliness_q3"  ,                                
        "misperception.about.the.world_q3" ,"political.conflict_q3" , "authoritarianism_q3", "stunted.child.development_q3","pessimism_q3",                                   
        "intimate.relations_q3" , "autobiographical.memory_q3") 

Q3.q <- c("Estrangement\nAlienation","Career Disruptions", "Low Qual. Education", "Low Trust\nin Science","Low Wellbeing"   ,                     
          "Economic Hardship" ,"Educational Inequality", "Erosion of Democratic Instit.", "Social Inequality","Mistrust",                            
          "Prejudice Racism", "Despair", "Self-Centeredness","Irrationality","Loneliness"  ,                                
          "Misperception of the World" ,"Political Conflict" , "Authoritarianism", "Stunted Child\nDevelopment","Pessimism",                                   
          "Intimate Relations" , "Autobiographical Memory")

Q3.type <- c("Interpersonal","Work", "Education", "Irrationality","Psych.\nWellbeing"   ,                     
             "Societal" ,"Education", "Societal", "Societal","Interpersonal",                            
             "Societal", "Psych.\nWellbeing", "Interpersonal","Irrationality","Psych.\nWellbeing"  ,                                
             "Irrationality" ,"Societal" , "Societal", "Education","Psych.\nWellbeing",                                   
             "Interpersonal" , "Psych.\nWellbeing")

Q3.data<-all.data %>% 
  summarise(across(estrangement...alienation_q3:autobiographical.memory_q3, mean, na.rm=TRUE))

colnames(Q3.data)[colnames(Q3.data)      # Rename two variable names
                     %in% Q3] <- Q3.q
Q3.data.long <- Q3.data %>% 
  pivot_longer(everything(),names_to = "Theme", values_to = "Proportion")
Q3.summ<-cbind(Q3.data.long, Q3.type)
colnames(Q3.summ) <- c("Theme","Proportion","Type")  # change column names
Q3.summ <- Q3.summ[order(Q3.summ$Proportion), ]  # sort
Q3.summ$Theme <- factor(Q3.summ$Theme, levels = Q3.summ$Theme)  # to retain the order in plot
Q3.summ$Percent<-Q3.summ$Proportion*100
Q3.plot<-ggplot(Q3.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() +
  geom_bar(stat="identity", width=.5) + ggtitle("Negative Consequences")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom")+labs(x="",y="%")
Q3.plot
```

## How many themes per person?

Did people just report 1-2 themes or a great number of themes? How do such trends vary across people?\
It turns out most people mentioned just one or two (median) negative consequences. Two people mentioned up 5 and 7 themes, respectively.\


```{r,echo=FALSE,warning=FALSE,message=FALSE}
all.data$Q3.pp<-rowSums(all.data[Q3])
ggplot(all.data, aes(x = Q3.pp)) +   geom_vline(xintercept=median(all.data$Q3.pp, na.rm=T), color="red")+
  geom_bar()+theme_pubclean()+scale_fill_d3()+labs(x="Number of Themes per Person",y="Count")
           
```


## Frequency Chart by Gender and Field

It is possible that socio-cultural experts focus *more* on social-structural issues than mental health and other possible domains of change because of idiosyncratic preferences.\
As the distribution of themes by field of expertise indicates, this is not likely to be the reason for mentioning political and interpersonal issues most (relative to mental health), with roughly similar distributions. The only exception is that sociocultural experts tended to emphasize mistrust, whereas non-SC experts emphasized irrationality/mispercetion of the world more. 

```{r,echo=FALSE,warning=FALSE,message=FALSE}
##BY gender
Q3.data.G<-all.data %>% group_by(Gender)%>% 
  summarise(across(estrangement...alienation_q3:autobiographical.memory_q3, mean, na.rm=TRUE))

colnames(Q3.data.G)[colnames(Q3.data.G)      # Rename two variable names
                    %in% Q3] <- Q3.q

Q3.data.G.long <- Q3.data.G %>% 
  pivot_longer(-Gender,names_to = "Theme", values_to = "Proportion")
Q3.G.summ<-cbind(Q3.data.G.long, Q3.type)
colnames(Q3.G.summ) <- c("Gender","Theme","Proportion","Type")  # change column names
Q3.G.summ$Percent<-Q3.G.summ$Proportion*100
Q3.G.plot<-ggplot(Q3.G.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Gender)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Negative Consequences")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom")+labs(x="",y="%")
Q3.G.plot

##BY socio-cultural vs. non-socio-cultural expertise
Q3.data.SC<-all.data %>% group_by(Field)%>% 
  summarise(across(estrangement...alienation_q3:autobiographical.memory_q3, mean, na.rm=TRUE))

colnames(Q3.data.SC)[colnames(Q3.data.SC)      # Rename two variable names
                     %in% Q3] <- Q3.q

Q3.data.SC.long <- Q3.data.SC %>% 
  pivot_longer(-Field,names_to = "Theme", values_to = "Proportion")
Q3.SC.summ<-cbind(Q3.data.SC.long, Q3.type)
colnames(Q3.SC.summ) <- c("Field","Theme","Proportion","Type")  # change column names
Q3.SC.summ$Percent<-Q3.SC.summ$Proportion*100
Q3.SC.plot<-ggplot(Q3.SC.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Field)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Negative Consequences")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom")+labs(x="",y="%")
Q3.SC.plot
```


## Frequencies by Cognitive Style: Dialectical Thinking and Consideration of Outsider Viewpoint

It appears that **people invoking dialecticism** in their reasoning were more likely to focus on *negative consequences (incl. socio-economic inequality, prejudice/racism, political conflict, authoritarianism)* as well as *low trust in science* in their negative forecasts as those who did not show dialecticism.\
\
It is hard to tell how similar **experts invoking baserate/outsider view information** were in their reasoning compared to experts who did not mention baserate info. There are some differences, but they may be due to differences in sub-sample size.\
Key differences that stand out are:\

 - greater focus on *low wellbeing*, *low trust in science*, and *estrangement/alienation* among *baseraters* and greater focus on *interpersinal mustruct* among *non-baseraters*.\ 

```{r,echo=FALSE,warning=FALSE,message=FALSE}
#Dialectical thinking

Q3.data.Dia<-all.data %>% group_by(Dialectic_Final)%>% 
  summarise(across(estrangement...alienation_q3:autobiographical.memory_q3, mean, na.rm=TRUE))

colnames(Q3.data.Dia)[colnames(Q3.data.Dia)      # Rename two variable names
                     %in% Q3] <- Q3.q
Q3.data.Dia$Dialecticism[Q3.data.Dia$Dialectic_Final==0]<-"Not Dialectical"
Q3.data.Dia$Dialecticism[Q3.data.Dia$Dialectic_Final==1]<-"Dialectical"

Q3.data.Dia.long <- Q3.data.Dia %>% 
  pivot_longer(-c("Dialectic_Final","Dialecticism"),names_to = "Theme", values_to = "Proportion")
Q3.Dia.summ<-cbind(Q3.data.Dia.long, Q3.type)
colnames(Q3.Dia.summ) <- c("Dialectical Code","Dialecticism","Theme","Proportion","Type")  # change column names
Q3.Dia.summ$Percent<-Q3.Dia.summ$Proportion*100
Q3.Dia.plot<-ggplot(Q3.Dia.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Dialecticism)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Negative Consequences")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="%")
Q3.Dia.plot

Q3.sum.Dia<-all.data %>% group_by(Dialectic_Final)%>% 
  summarise(across(estrangement...alienation_q3:autobiographical.memory_q3, sum, na.rm=TRUE))

#test significance of the effect - prevalence of themes in the dialectical vs. non-dialectical samples.
Q3.data.Dia.long<-all.data %>% pivot_longer(cols=estrangement...alienation_q3:autobiographical.memory_q3, names_to="Theme",values_to="Presence")

Q3.dia.result<-glmer(Presence ~ Theme*Dialectic_Final + (1 | Name), data = Q3.data.Dia.long, family = binomial)
summary(Q3.dia.result)
Anova(Q3.dia.result,type="III")


#outsider Viewpoint

Q3.data.Outsider<-all.data %>% group_by(Q3_Outside_Final)%>% 
  summarise(across(estrangement...alienation_q3:autobiographical.memory_q3, mean, na.rm=TRUE))

colnames(Q3.data.Outsider)[colnames(Q3.data.Outsider)      # Rename two variable names
                     %in% Q3] <- Q3.q
Q3.data.Outsider$OutsiderV[Q3.data.Outsider$Q3_Outside_Final==0]<-"No Baserate Info (25%)"
Q3.data.Outsider$OutsiderV[Q3.data.Outsider$Q3_Outside_Final==1]<-"Mentioned Baserates (75%)"

Q3.data.Outsider.long <- Q3.data.Outsider %>% drop_na(OutsiderV)%>% 
  pivot_longer(-c("Q3_Outside_Final","OutsiderV"),names_to = "Theme", values_to = "Proportion")
Q3.Outsider.summ<-cbind(Q3.data.Outsider.long, Q3.type)
colnames(Q3.Outsider.summ) <- c("Outsider Code","Baserates","Theme","Proportion","Type")  # change column names
Q3.Outsider.summ$Percent<-Q3.Outsider.summ$Proportion*100
Q3.Outsider.plot<-ggplot(Q3.Outsider.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Baserates)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Negative Consequences")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="%")
Q3.Outsider.plot


```


## PCA 

```{r,echo=FALSE,warning=FALSE,message=FALSE}

vss(all.data[Q3], rotate="oblimin", fm="pc")
principal(all.data[Q3], nfactors = 8, rotate="oblimin")


```

Similar to positive consequences, we can address the question of diversity by examining the degree to which scores across themes are reducible to common component, examining principal component analysis (PCA).\
It appears that 7 or 8 components is as good a start to account for 22 themes for Question 3 (neg consequences), as any other. This said, the scree plot line is pretty flat.\
Again, keep in mind that reducing 22 items to 8 components is not much. Moreover, when we reduce the items to 8 components, the first component explains only 10 % of the variance and in total 8 components account only for 64%.\
Given that each theme by default explains 4.6% of the variance (1/22), this is not much. \
Furthermore, each of these components is largely based on 1-2 items (loadings > .6), with an exception of TC1 - decline in education, stunted child dev & intomate relationships \

In short, principal components show substantial diversity, with the top factors political conflict and prejudice/racism not accounting for more than 16% of the variance and showing only weak intercorrelation among themselves\

## Network Graph

```{r,echo=FALSE,warning=FALSE,message=FALSE}
Q3.cor<-partial.r(all.data,Q3,cs(tokens.Q3),method="spearman")

row.names(Q3.cor) <-Q3.q
colnames(Q3.cor) <-Q3.q

# Keep only meaningful correlations (<.17)
Q3.cor[Q3.cor<0.17] <- 0

# Make an Igraph object from this matrix:
Q3.network <- graph_from_adjacency_matrix( Q3.cor, weighted=T, mode="undirected", diag=F)



#color palette
coul <- brewer.pal(nlevels(as.factor(Q3.type)), "Dark2")

# Map the color to types
my_color <- coul[as.numeric(as.factor(Q3.type))]

# plot with edges

plot(Q3.network,
     edge.color=rep(c("red","pink"),5),           # Edge color
     edge.width=seq(1,10),                        # Edge width, defaults to 1
     edge.arrow.size=1,                           # Arrow size, defaults to 1
     edge.arrow.width=1,                          # Arrow width, defaults to 1
     edge.lty=c("solid"),
     vertex.label.color="black", 
                vertex.color=my_color                           # Line type, could be 0 or "blank", 1 or "solid", 2 or "dashed", 3 or "dotted", 4 or "dotdash", 5 or "longdash", 6 or "twodash"
     #edge.curved=c(rep(0,5), rep(1,5))            # Edge curvature, range 0-1 (FALSE sets it to 0, TRUE to 0.5)
)


```

```{r,echo=FALSE,warning=FALSE,message=FALSE}

Q3.ceb <- cluster_edge_betweenness(Q3.network) 

dendPlot(Q3.ceb, mode="hclust")
plot(Q3.ceb, Q3.network, axes=T,edge.width=seq(1,10)) 


```

As we can see above, even after removing negligible correlations (r < .17), cluster analyses on top of the network graphs show four clusters - Four clusters: irrationality & misperception of the world, prejudice, & decline of democracy / well-being decline and associated factors / family matters & educational (social) inequality (incl stunted child dev) / economic hardships.\

Themes for negative consequences of the pandemic appear largely distinct, except for link between loneliness and gratitude, linking two networks concerning family matters (loneliness) and wellbeing (alienation/estrangement).\

```{r,echo=FALSE,warning=FALSE,message=FALSE}
edgebundle(Q3.network, padding = 120, fontsize = 10)

```


## Convergence vs. divergence of themes over time

As a reminder: I started to interview people in late June, 2020, in the weeks following the BLM protests and riots following George Floyd's death. Over the course of the summer and the fall, numerous other events occurred, including relaxing and re-introducing lockdown policies and regulations in different parts of the world, start of in-person schooling in the fall, the US Presidential election, and its aftermath.\
Once again, we binned scores in the the same four groups as above: June, July, Sept/early Oct, and second part of Oct-early Dec.\


```{r,echo=FALSE,warning=FALSE,message=FALSE}

Q3.data.M<-all.data %>% group_by(M)%>%
  summarise(across(estrangement...alienation_q3:autobiographical.memory_q3, sum,na.rm=T))


colnames(Q3.data.M)[colnames(Q3.data.M)      # Rename two variable names
                     %in% Q3] <- Q3.q

Q3.data.long.M <- Q3.data.M %>% 
  pivot_longer(-M,names_to = "Theme", values_to = "Count")
Q3.summ.M<-cbind(Q3.data.long.M, Q3.type)
colnames(Q3.summ.M) <- c("Period","Theme","Count","Type")  # change column names
Q3.summ.M <- Q3.summ.M[order(Q3.summ.M$Count), ]  # sort

Q3.data.T<-Q3.summ.M %>% group_by(Period, Type)%>%
    summarise(score=sum(Count))

countByMonth<-all.data %>% group_by(M)%>% 
  summarise(count = n() )

colnames(countByMonth) <- c("Period","N")  # change column names

Q3.data.TT<-full_join(Q3.data.T,countByMonth, by = "Period")
Q3.data.TT$Percent<-(Q3.data.TT$score/Q3.data.TT$N)*100

ggplot(Q3.data.TT, aes(x=Period, y=Percent, group = Type, color=Type, fill=Type)) + 
    geom_line( ) +
    geom_point(shape=21,  size=4) +
   theme_pubclean()+scale_color_d3()+scale_fill_d3()


```

Focus on societal issues consistently remains the most frequent topic over time.\
Whereas discussion of health and well-being was the second most frequent category mentioned for negative consequences in June, it become third/forth frequent from July on.\
In exchange, irrationality started to increase in frequency being mentioned in the fall, closer to the US Election and second lockdowns.\
Least frequently mentioned were concerns about work.

Like for positive consequences, one can post-hoc speculate that some of the major themes such as well-being related concerns (e.g., loneliness, well-being decline), and forecasts of irrationality (incl. mispeception of the world, mistrust in science) were likely impacted by societal events unfolding around the time people were making forecasts. \

Notably, one theme remained pretty stable over time, with negative societal consequences (political conflict, prejudice and racism, societal inequality) - the most frequently mentioned topic in response to this question - remained consistently most frequent over time.\

# Comparison of Positive and Negative Consequences Side by Side

```{r,echo=FALSE,warning=FALSE,message=FALSE}

Q1.summ <- Q1.summ[order(Q1.summ$Percent), ]  # sort
Q1.summ$Theme <- factor(Q1.summ$Theme, levels = Q1.summ$Theme)  # to retain the order in plot

Q3.summ <- Q3.summ[order(Q3.summ$Percent), ]  # sort
Q3.summ$Theme <- factor(Q3.summ$Theme, levels = Q3.summ$Theme)  # to retain the order in plot


#prepare data for lollipop chart
Q1.summ2<- Q1.summ[order(Q1.summ$Percent), ]
Q1.summ3<-Q1.summ2%>%
  mutate(Theme = factor(Theme, levels = .$Theme))
Q3.summ2<-Q3.summ%>%
  mutate(Theme = factor(Theme, levels = .$Theme))


data.Q1 <- data.frame(
  Theme=Q1.summ3$Theme,
  Frequency=Q1.summ3$Percent,
  group=Q1.summ3$Type
)

data.Q3<-data.frame(
  Theme=Q3.summ2$Theme,
  Frequency=-Q3.summ2$Percent,
  group=Q3.summ2$Type
)

data.Q13<-rbind(data.Q1,data.Q3)
data.Q13<- data.Q13[order(data.Q13$Frequency), ]
data.Q13<-data.Q13%>%
  arrange(Frequency)%>%
  mutate(Above = ifelse(Frequency > 0, TRUE, FALSE),
         Theme = factor(Theme, levels = .$Theme))
# plot Lollipop chart
fig<-ggplot(data.Q13, aes(x=Theme, y=Frequency,color = Above, tooltip=group))+
   geom_segment(aes(x=Theme, xend=Theme, y=0, yend=Frequency), color = "burlywood3", size = 0.6) +
  geom_point(size=3)+
  #geom_text(color = 'black', size = 2)+
   scale_y_continuous(breaks=c(-20,-10,0,10,25),
        labels=c("20%", "10%", "0","10%","25%"))+
   coord_flip() +scale_color_d3()+theme_pubclean()+
  xlab("Themes")+geom_hline(yintercept = 0, linetype="dashed", 
                color = "black", size=0.7)+ylab("")+
    theme(legend.title=element_blank(), legend.position="none",legend.text=element_text(size=5))
fig
ggplotly(fig, height = 920)


#get a graph with top themes

data.Q1x <-tail(data.Q1,10)

data.Q3x <-tail(data.Q3,10)

data.Q13x<-rbind(data.Q1x,data.Q3x)
data.Q13x<- data.Q13x[order(data.Q13x$Frequency), ]
data.Q13x<-data.Q13x%>%
  arrange(Frequency)%>%
  mutate(Above = ifelse(Frequency > 0, TRUE, FALSE),
         Theme = factor(Theme, levels = .$Theme))
# plot Lollipop chart
fig13x<-ggplot(data.Q13x, aes(x=Theme, y=Frequency,color=Above))+
  geom_segment(aes(x=Theme, xend=Theme, y=0, yend=Frequency), color = "burlywood3", size = 0.6) +
  geom_point(size=4)+
  #geom_text(color = 'black', size = 2)+
  scale_y_continuous(breaks=c(-20,-10,0,10,20),
        labels=c("20%", "10%", "0","10%","20%")) +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    axis.ticks.x = element_blank()
  ) +scale_color_tableau() +
  xlab("Themes")+geom_hline(yintercept = 0, linetype="dashed", 
                color = "black", size=0.7)+ylab("")+
  theme(legend.title = element_blank())+
  theme(legend.position="none",legend.text=element_text(size=7)) +
  scale_color_manual(labels = c("Negative Consequences", "Positive Consequences"),
                     values = c("#4E79A7","#F28E2B"))+theme(panel.background = element_rect(fill = "#f3f2f0"),plot.background = element_rect(fill = "#f3f2f0"),legend.background = element_rect(fill = "#f3f2f0"))
fig13x


```


# Change in Top Consequences over Time

We examined whether top positive and negative categorized changed over the course of the pandemic. To this end, we focused on the most frequent six themes mentioned across positive and negative forecasts.\
Experts were more likely to predict greater will for political and structural societal change, as well as greater prejudice and racism after George Floyd death and anti-police brutality protests in early summer. When approaching new lockdowns in the fall of 2020, topics such as social inequality became more dominant in expert reflections. And in the week preceding and following the highly polarized 2020 US Presidential election in early November, solidarity and political conflict entered the centerstage. Because experts were explicitly instructed to provide forecasts for the timeframe of several years after the pandemic, the event-contingent fluctuation in forecast could reflect focalism in expert predictions (Wilson, Wheatly, Meyers, Gilbert, & Axsom, 2000) or Bayesian information updating based on pressing societal events of the moment. Regardless of possible reasons, the cross-temporal variability in forecasts emphasizes cross-temporal uncertainty in expert predictions.\

```{r,echo=FALSE,warning=FALSE,message=FALSE}
top.cat.5<-c("social.connectedness_q1","political.engagement...structural.change_q1","solidarity_q1","reconsidering.habits_q1","embrace.new.tech_q1","political.conflict_q3","prejudice...racism_q3","exacerbated.social.inequality_q3","mistrust_q3","decreased.well.being_q3")
top.cat<-c("social.connectedness_q1","political.engagement...structural.change_q1","solidarity_q1","political.conflict_q3","prejudice...racism_q3","exacerbated.social.inequality_q3")

#perform logistic regression. to test whether there is a difference by month.

temporal.data<-all.data %>% pivot_longer(cols=top.cat, names_to="theme",values_to="count")
temporal.data$Theme[temporal.data$theme=="social.connectedness_q1"]<-"Social\nConnectedness"
temporal.data$Theme[temporal.data$theme=="political.engagement...structural.change_q1"]<-"Polit.\nStructure-Change"
temporal.data$Theme[temporal.data$theme=="solidarity_q1"]<-"Solidarity"
temporal.data$Theme[temporal.data$theme=="political.conflict_q3"]<-"Political Conflict"
temporal.data$Theme[temporal.data$theme=="prejudice...racism_q3"]<-"Prejudice Racism"
temporal.data$Theme[temporal.data$theme=="exacerbated.social.inequality_q3"]<-"Social Inequality"

temporal.data$pos[temporal.data$theme=="social.connectedness_q1"]<-1
temporal.data$pos[temporal.data$theme=="political.engagement...structural.change_q1"]<-1
temporal.data$pos[temporal.data$theme=="solidarity_q1"]<-1
temporal.data$pos[temporal.data$theme=="political.conflict_q3"]<-0
temporal.data$pos[temporal.data$theme=="prejudice...racism_q3"]<-0
temporal.data$pos[temporal.data$theme=="exacerbated.social.inequality_q3"]<-0

temp.result.pos<-glmer(count ~ week*Theme + (1 | Name), data = subset(temporal.data,pos==1), family = binomial)
Anova(temp.result.pos,type="III")
pos.pl<-sjPlot::plot_model(temp.result.pos,type="int")

temp.result.neg<-glmer(count ~ week*Theme + (1 | Name), data = subset(temporal.data,pos==0), family = binomial)
Anova(temp.result.neg,type="III")
neg.pl<-sjPlot::plot_model(temp.result.neg,type="int")

temp.result.pos.emt<-emmeans::emtrends(temp.result.pos,"Theme",var="week")
pairs(temp.result.pos.emt)

temp.result.neg.emt<-emmeans::emtrends(temp.result.neg,"Theme",var="week")
pairs(temp.result.neg.emt)

pos.plot<-pos.pl+scale_x_continuous(breaks=c(25,30,37,42,49),labels=c("15 June", "20 July", "9 Sept","12 Oct","30 Nov"))+ylim(NA,.8)+scale_y_continuous(breaks=c(0,.2,.4,.6,.8),
        labels=c("0%", "20%", "40%","60%","80%"))+
   theme_pubclean()+scale_color_colorblind()+scale_fill_colorblind()+theme(legend.title=element_blank())+ylab("Theme's Likelihood")+xlab("")+ggtitle("")
neg.plot<-neg.pl+scale_x_continuous(breaks=c(25,30,37,42,49),labels=c("15 June", "20 July", "9 Sept","12 Oct","30 Nov"))+ylim(NA,.8)+scale_y_continuous(breaks=c(0,.2,.4,.6,.8),
         labels=c("0%", "20%", "40%","60%","80%"))+
   theme_pubclean()+scale_color_colorblind()+scale_fill_colorblind()+theme(legend.title=element_blank())+ylab("Theme's Likelihood")+xlab("")+ggtitle("")

ggarrange(pos.plot,neg.plot,ncol=1,label.x="Month")

##raw score version (proportion for different periods)

temporal.change<-all.data %>% group_by(M)%>%
  summarise(across(top.cat, sum,na.rm=T))

top.cat.names.5<-c("Period","Social\nConnectedness","Polit.\nStructure-Change","Solidarity","Reconsider Habits","Embrace New Tech","Political Conflict","Prejudice Racism","Social Inequality","Mistrust","Low Wellbeing")
temporal.change.type.5<-c("Positive","Positive","Positive","Positive","Positive","Negative","Negative","Negative","Negative","Negative")

top.cat.names<-c("Period","Social\nConnectedness","Polit.\nStructure-Change","Solidarity","Political Conflict","Prejudice Racism","Social Inequality")
temporal.change.type<-c("Positive","Positive","Positive","Negative","Negative","Negative")

colnames(temporal.change) <- top.cat.names
temporal.change.long <- temporal.change %>% 
  pivot_longer(-Period,names_to = "Theme", values_to = "Count")

temporal.change.long<-cbind(temporal.change.long, temporal.change.type)
colnames(temporal.change.long) <- c("Period","Theme","Count","Type")  # change column names
temporal.change.long <- temporal.change.long[order(temporal.change.long$Count), ]  # sort

countByMonth<-all.data %>% group_by(M)%>% 
  summarise(count = n() )

colnames(countByMonth) <- c("Period","N")  # change column names

temporal.change.TT<-full_join(temporal.change.long,countByMonth, by = "Period")
temporal.change.TT$Percent<-(temporal.change.TT$Count/temporal.change.TT$N)*100

temp.change.graph<-ggplot(temporal.change.TT, aes(x=Period, y=Percent, group = Theme, color=Theme, fill=Theme)) + 
    geom_line( ) +
    geom_point(shape=21,  size=4) +facet_wrap(~Type, ncol=1,as.table = FALSE)+
   theme_pubclean()+scale_color_d3()+scale_fill_d3()+theme(legend.title=element_blank())
temp.change.graph

# does it matter if one is from the US or not for type of trajectories?

temp.result.pos.location<-glmer(count ~ week*Theme*Location + (1 | Name), data = subset(temporal.data,pos==1), family = binomial)
Anova(temp.result.pos.location,type="III")
pos.pl.location<-sjPlot::plot_model(temp.result.pos.location,type="int")

temp.result.neg.location<-glmer(count ~ week*Theme*Location+ (1 | Name), data = subset(temporal.data,pos==0), family = binomial)
Anova(temp.result.neg.location,type="III")
neg.pl.location<-sjPlot::plot_model(temp.result.neg.location,type="int")



pos.pl.location<-pos.pl.location[[4]]+scale_x_continuous(breaks=c(25,30,37,42,49),labels=c("15 June", "20 July", "9 Sept","12 Oct","30 Nov"))+ylim(NA,.8)+scale_y_continuous(breaks=c(0,.2,.4,.6,.8),labels=c("0%", "20%", "40%","60%","80%"))+
   theme_pubclean()+scale_color_colorblind()+scale_fill_colorblind()+theme(legend.title=element_blank())+ylab("Theme's Likelihood")+xlab("")+ggtitle("")
neg.pl.location<-neg.pl.location[[4]]+scale_x_continuous(breaks=c(25,30,37,42,49),labels=c("15 June", "20 July", "9 Sept","12 Oct","30 Nov"))+ylim(NA,.8)+scale_y_continuous(breaks=c(0,.2,.4,.6,.8),
         labels=c("0%", "20%", "40%","60%","80%"))+
   theme_pubclean()+scale_color_colorblind()+scale_fill_colorblind()+theme(legend.title=element_blank())+ylab("Theme's Likelihood")+xlab("")+ggtitle("")

ggarrange(pos.pl.location,neg.pl.location,ncol=1,label.x="Month")

```
# Wisdom for Positive Consequences

## Summary

We identified 21 distinct categories. 
Most themes mentioned by less than 10% of the interviewees.\
Only one theme was mentioned by at least ten people: Need for solidarity.\

Most common wisdom-related strategies to sustain positive changes (7+ ppl) concern solidarity, willingness to engage in political and structure change, perspective-taking, and critical thinking. \

## Frequency Chart

```{r,echo=FALSE,warning=FALSE,message=FALSE}
##'QUESTION 2

Q2 <- c("social.connectedness_q2", "social.support_q2",  "evidence.based.decision.making_q2","improved.work.life.balance_q2","acknowledge.of.uncertainty.flexibility_q2","balance.of.personal...others.interest_q2","perspective.taking_q2",
        "critical.thinking_q2","bipartisanship.and.international.cooperation_q2","solidarity_q2",                        
        "improve.communication_q2", "intellectual.humility_q2","living.in.the.moment_q2",
        "learning.from.pandemics_q2","political.engagement...structural.change_q2",
        "self.reflection.on.what.s.important_q2","shared.humanity_q2","sympathy...compassion_q2", "self.distancing_q2","embrace.new.tech_q2","personal.resilience_q2") 


Q2.q <- c("Social Connectedness",  "Social Support", "Evidence-based Judgement", "Work-Life Balance", "Acknowledge Uncertainty","Balance Diverse Interests","Perspective-taking",
          "Critical Thinking","Polit.Cooperation",
          "Solidarity", "Improved Communication", "Intellectual Humility","Live in the Moment",
          "Learning from Pandemics","Polit.Structural Change","What's Important?","Shared Humanity","Sympathy&Compassion", "Self-distancing","Embrace New Tech","Personal Resilience") 

Q2.type <- c("Prosocial","Prosocial","Quest\nfor Knowledge","Work",        "Meta-cognition","Meta-cognition","Meta-cognition",
             "Meta-cognition","Societal","Prosocial",                        
             "Interpersonal", "Meta-cognition","Psych.\nWellbeing",
             "Quest\nfor Knowledge","Societal",
             "Insight","Meta-cognition","Meta-cognition", "Meta-cognition","Work","Psych.\nWellbeing") 


Q2.data<-all.data %>% 
  summarise(across(social.connectedness_q2:personal.resilience_q2, mean, na.rm = T))




colnames(Q2.data)[colnames(Q2.data)      # Rename two variable names
                     %in% Q2] <- Q2.q

Q2.data.long <- Q2.data %>% 
  pivot_longer(everything(),names_to = "Theme", values_to = "Proportion")
Q2.summ<-cbind(Q2.data.long, Q2.type)
colnames(Q2.summ) <- c("Theme","Proportion","Type")  # change column names
Q2.summ <- Q2.summ[order(Q2.summ$Proportion), ]  # sort
Q2.summ$Theme <- factor(Q2.summ$Theme, levels = Q2.summ$Theme)  # to retain the order in plot
Q2.summ$Percent<-Q2.summ$Proportion*100
Q2.plot<-ggplot(Q2.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() +
  geom_bar(stat="identity", width=.5) + ggtitle("Wisdom for Positive Consequences")+theme_pubclean()+scale_fill_d3()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom", legend.text=element_text(size=6.5))+labs(x="",y="%")
Q2.plot
```

## How many themes per person?

Did people just report 1-2 themes or a great number of themes? How do such trends vary across people?\
It turns out most people mentioned just one or two (median) negative consequences. Two people mentioned up to 4 and 5 themes, respectively.\


```{r,echo=FALSE,warning=FALSE,message=FALSE}
all.data$Q2.pp<-rowSums(all.data[Q2])
ggplot(all.data, aes(x = Q2.pp)) +   geom_vline(xintercept=median(all.data$Q2.pp, na.rm=T), color="red")+
  geom_bar()+theme_pubclean()+scale_fill_d3()+labs(x="Number of Themes per Person",y="Count")
           
```

## Frequencies by Gender and Field

It is not surprising that wisdom experts would mention meta-cognitive and moral considerations. The question is whether scientists without much familiarity with the science of wisdom would mention similar constructs. It appears that is the case, with non-wisdom experts emphasizing such meta-cognitive categories as critical thinking, intellectual humility, acknowledgment of uncertainty, sympathy and compassion and perspective-taking, along with moral aspirations about solidarity, social support, and paying greater attention to one's family and relationships, and promoting societal change toward fair and just society.

```{r,echo=FALSE,warning=FALSE,message=FALSE}
##BY gender
Q2.data.G<-all.data %>% group_by(Gender)%>% 
  summarise(across(social.connectedness_q2:personal.resilience_q2, mean, na.rm = TRUE))

colnames(Q2.data.G)[colnames(Q2.data.G)      # Rename two variable names
                    %in% Q2] <- Q2.q

Q2.data.G.long <- Q2.data.G %>% 
  pivot_longer(-Gender,names_to = "Theme", values_to = "Proportion")
Q2.G.summ<-cbind(Q2.data.G.long, Q2.type)
colnames(Q2.G.summ) <- c("Gender","Theme","Proportion","Type")  # change column names
Q2.G.summ$Percent<-Q2.G.summ$Proportion*100
Q2.G.plot<-ggplot(Q2.G.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Gender)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Wisdom for Positive Consequences")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom", legend.text=element_text(size=6.5))+labs(x="",y="%")

Q2.G.plot

##BY wisdom familiarity
Q2.data.W<-all.data %>% group_by(FamiliarWisdom)%>% 
  summarise(across(social.connectedness_q2:personal.resilience_q2, mean, na.rm = TRUE))

colnames(Q2.data.W)[colnames(Q2.data.W)      # Rename two variable names
                     %in% Q2] <- Q2.q

Q2.data.W.long <- Q2.data.W %>% 
  pivot_longer(-FamiliarWisdom,names_to = "Theme", values_to = "Proportion")
Q2.W.summ<-cbind(Q2.data.W.long, Q2.type)
colnames(Q2.W.summ) <- c("FamiliarWisdom","Theme","Proportion","Type")  # change column names
Q2.W.summ$Percent<-Q2.W.summ$Proportion*100
Q2.W.plot<-ggplot(Q2.W.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ FamiliarWisdom)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Wisdom for Positive Consequences\nby Wisdom Familarity")+theme_pubclean()+scale_fill_d3()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom", legend.text=element_text(size=6.5))+labs(x="",y="%")
Q2.W.plot


#test whether familiarity with wisdom impacts type of themes selected
Q2.data.Famil.long<-all.data %>% pivot_longer(cols=social.connectedness_q2:personal.resilience_q2, names_to="Theme",values_to="Presence")

Q2.Famil.result<-glmer(Presence ~ Theme*FamiliarWisdom + (1 | Name), data = Q2.data.Famil.long, family = binomial)
summary(Q2.Famil.result)
Anova(Q2.Famil.result,type="III")
#no significant difference by familiarity with wisdom scholarship

##BY socio-cultural vs. non-socio-cultural expertise
Q2.data.SC<-all.data %>% group_by(Field)%>% 
  summarise(across(social.connectedness_q2:personal.resilience_q2, mean, na.rm = TRUE))

colnames(Q2.data.SC)[colnames(Q2.data.SC)      # Rename two variable names
                     %in% Q2] <- Q2.q

Q2.data.SC.long <- Q2.data.SC %>% 
  pivot_longer(-Field,names_to = "Theme", values_to = "Proportion")
Q2.SC.summ<-cbind(Q2.data.SC.long, Q2.type)
colnames(Q2.SC.summ) <- c("Field","Theme","Proportion","Type")  # change column names
Q2.SC.summ$Percent<-Q2.SC.summ$Proportion*100
Q2.SC.plot<-ggplot(Q2.SC.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Field)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Wisdom for Positive Consequences")+theme_pubclean()+scale_fill_d3()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom", legend.text=element_text(size=6.5))+labs(x="",y="%")
Q2.SC.plot

Q2.SC.result<-glmer(Presence ~ Theme*Field + (1 | Name), data = Q2.data.Famil.long, family = binomial)
summary(Q2.SC.result)
Anova(Q2.SC.result,type="III")

```


## Frequencies by Cognitive Style: Dialectical Thinking

### Is there a difference in themes mentioned between experts who show dialecticism in their reasoning about societal change vs. those who do not invoke dialectism?

It appears that **people invoking dialecticism** in their reasoning were more likely to raise the themes of *perspective-taking*, *intellectual humility*, *critical thinking*,*social connectedness*, *political cooperation*, and *work-life balance*  compared to those who did not show dialecticism.\
\

```{r,echo=FALSE,warning=FALSE,message=FALSE}
#Dialectical thinking

Q2.data.Dia<-all.data %>% group_by(Dialectic_Final)%>% 
  summarise(across(social.connectedness_q2:personal.resilience_q2, mean, na.rm=TRUE))

colnames(Q2.data.Dia)[colnames(Q2.data.Dia)      # Rename two variable names
                     %in% Q2] <- Q2.q
Q2.data.Dia$Dialecticism[Q2.data.Dia$Dialectic_Final==0]<-"Not Dialectical"
Q2.data.Dia$Dialecticism[Q2.data.Dia$Dialectic_Final==1]<-"Dialectical"

Q2.data.Dia.long <- Q2.data.Dia %>% 
  pivot_longer(-c("Dialectic_Final","Dialecticism"),names_to = "Theme", values_to = "Proportion")
Q2.Dia.summ<-cbind(Q2.data.Dia.long, Q2.type)
colnames(Q2.Dia.summ) <- c("Dialectical Code","Dialecticism","Theme","Proportion","Type")  # change column names
Q2.Dia.summ$Percent<-Q2.Dia.summ$Proportion*100
Q2.Dia.plot<-ggplot(Q2.Dia.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Dialecticism)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Wisdom for Positive Consequences by Cognitive Style")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="%")
Q2.Dia.plot


```

### Is there a difference in **Common Wisdom Model** themes between dialectical experts vs. those who do not invoke dialectism?

It appears there is a significant difference in overall prevalence of CWM components (morals and meta-cog) over other categories. Further, there is a trend toward this difference being more pronounced among folks reasoning in a dialectical fashion about their forecasts (z-score ratio of 1.61 to 1: 4.08 vs. 2.53).


```{r,warning=FALSE,message=FALSE}
#Are dialectical thinkers more likely to mention core aspects of wisdom as identified by the Common wisdom model per Toronto Wisdom Task Force 2019?

#sum up all morals
morals.Q2<-c("social.connectedness_q2", "social.support_q2","bipartisanship.and.international.cooperation_q2","solidarity_q2",                        
        "political.engagement...structural.change_q2","shared.humanity_q2") 
        
#sum up all metacognitive
metacog.Q2<- c("acknowledge.of.uncertainty.flexibility_q2","balance.of.personal...others.interest_q2","perspective.taking_q2",
        "critical.thinking_q2","intellectual.humility_q2","sympathy...compassion_q2", "self.distancing_q2") 
#sum the rest
others.Q2<-c("evidence.based.decision.making_q2","improved.work.life.balance_q2",
             "improve.communication_q2", "living.in.the.moment_q2",
        "learning.from.pandemics_q2","self.reflection.on.what.s.important_q2","embrace.new.tech_q2","personal.resilience_q2")

all.data$morals.Q2<-rowSums(all.data[morals.Q2])
all.data$metacog.Q2<-rowSums(all.data[metacog.Q2])
all.data$others.Q2<-rowSums(all.data[others.Q2])
all.data$CWM.Q2<-rowSums(all.data[c("morals.Q2","metacog.Q2")])

all.data$Dialecticism[all.data$Dialectic_Final==0]<-"Not Dialectical"
all.data$Dialecticism[all.data$Dialectic_Final==1]<-"Dialectical"

Q2.dia.data<-all.data %>% pivot_longer(cols=morals.Q2:others.Q2, names_to="type",values_to="count")

Q2.dia.result<-glmer(count ~ type*Dialecticism + (1 | Name), data = Q2.dia.data, family = poisson)
Anova(Q2.dia.result,type="III") #no significant differences
sjPlot::plot_model(Q2.dia.result,type="int")

Q2.dia.result.emt<-emmeans::emmeans(Q2.dia.result,"type",var="Dialecticism")
pairs(Q2.dia.result.emt)

#contrast of cWM (morals and meta-cog) vs. the rest

Q2.dia.data.C<-all.data %>% pivot_longer(cols=c("others.Q2","CWM.Q2"), names_to="type",values_to="count")

Q2.dia.result.C<-glmer(count ~ type*Dialecticism + (1 | Name), data = Q2.dia.data.C, family = poisson)
Anova(Q2.dia.result.C,type="III") #no significant differences
sjPlot::plot_model(Q2.dia.result.C,type="int")

Q2.dia.result.C.emt<-emmeans::emmeans(Q2.dia.result.C,"type",by="Dialecticism")
pairs(Q2.dia.result.C.emt)



#look at it in terms of raw counts
Q2.data.Dia.CWM<-all.data %>% group_by(Dialecticism)%>% 
  summarise(across(morals.Q2:others.Q2, sum, na.rm=TRUE))%>%
pivot_longer(-Dialecticism,names_to = "Theme", values_to = "Count")

Q2.Dia.CWM.plot<- ggplot(Q2.data.Dia.CWM,aes(x=Dialecticism,y=Count,fill=Theme))+
  geom_col(stat="identity")+theme_pubclean()+scale_fill_d3()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="Count")

Q2.Dia.CWM.plot #more meta-cognition and more morals among dialectics

mosaic(xtabs(Count~ Dialecticism+Theme, data= Q2.data.Dia.CWM), shade=TRUE) #but not significant


Q2.data.W.CWM<-all.data %>% group_by(FamiliarWisdom)%>% 
  summarise(across(morals.Q2:others.Q2, sum, na.rm=TRUE))%>%
pivot_longer(-FamiliarWisdom,names_to = "Theme", values_to = "Count")
mosaic(xtabs(Count~ FamiliarWisdom+Theme, data= Q2.data.W.CWM), shade=TRUE) #those familiar with wisdom research are more likely to mention meta-cognition, but not significantly so.

chisq.test(xtabs(Count~ FamiliarWisdom+Theme, data= Q2.data.W.CWM))

```

## PCA 

```{r,echo=FALSE,warning=FALSE,message=FALSE}

vss(all.data[Q2], rotate="oblimin", fm="pc")
principal(all.data[Q2], nfactors = 8, rotate="oblimin")


```

Similar to prior results, we can address the question of diversity by examining the degree to which scores across themes are reducible to common component, examining principal component analysis (PCA).\
It appears that 7 or 8 components is as good a start to account for 21 themes for Question 2 (wisdom for pos consequences), as any other. Once again, the scree plot line is pretty flat.\
Again, keep in mind that reducing 21 items to 8 components is not parsimonious. Moreover, when we reduce the items to 8 components, the first component explains only 11 % of the variance, and in total 8 components explain 61% of variance.\
Given that each theme by default explains 4.8% of the variance (1/21), this is not much. \
Furthermore, each of these components is largely based on 1-2 items (loadings > .6), with an exception of TC1 - social connectedness, social support, and balance of personal and others interests, and improved work-life balance. \

In short, principal components show substantial diversity, with the top factor solidarity not even loading squarely on any of the components.\



## Network Graph

Even after removing negligible correlations (r < .17), cluster analyses on top of the network graphs show seven clusters, two of which are strongly inter-related (social connectedness and improved communication): \
•	Critical thinking/intellectual humility \
•	Willingness to adopt/learn new tech \
•	Meta-cognition: perspective-taking, balancing different interests, acknowledge uncertainty, and resilience \
•	Social cognitions: Balance of different interests, social connectedness, sympathy, social support \
•	Big picture focus: Shared humanity and self-distancing \
•	Solidarity, political engagement & cooperation \
•	Focus on what's important/mindfulness of the present \
•	Key is social connectedness - it connects two clusters. Solidarity also appears critical and close to the center of the network (proximity to most themes) \

```{r,echo=FALSE,warning=FALSE,message=FALSE}
#create correlation matrix, adjusting for number of words in each narrative
Q2.cor<-partial.r(all.data,Q2,cs(tokens.Q2),method="spearman")
row.names(Q2.cor) <-Q2.q
colnames(Q2.cor) <-Q2.q

# Keep only meaningful correlations (<.19)
Q2.cor[Q2.cor<0.17] <- 0

# Make an Igraph object from this matrix:
Q2.network <- graph_from_adjacency_matrix( Q2.cor, weighted=T,  mode="undirected", diag=F)


#color palette
coul <- brewer.pal(nlevels(as.factor(Q2.type)), "Dark2")

# Map the color to types
my_color <- coul[as.numeric(as.factor(Q2.type))]

# plot with edges

plot(Q2.network,
     edge.color=rep(c("red","pink"),5),           # Edge color
     edge.width=seq(1,10),                        # Edge width, defaults to 1
     edge.arrow.size=1,                           # Arrow size, defaults to 1
     edge.arrow.width=1,                          # Arrow width, defaults to 1
     edge.lty=c("solid"),
     vertex.label.color="black", 
     vertex.color=my_color# Line type, could be 0 or "blank", 1 or "solid", 2 or "dashed", 3 or "dotted", 4 or "dotdash", 5 or "longdash", 6 or "twodash"
     #edge.curved=c(rep(0,5), rep(1,5))            # Edge curvature, range 0-1 (FALSE sets it to 0, TRUE to 0.5)
)



Q2.ceb <- cluster_edge_betweenness(Q2.network) 

dendPlot(Q2.ceb, mode="hclust")
edge.start <- ends(Q2.network, es=E(Q2.network), names=F)[,1]
edge.col <- V(Q2.network)$color[edge.start]

plot(Q2.ceb, Q2.network, axes=F, ,vertex.shape="none",edge.color=edge.col, edge.curved=.1,
     vertex.label.color="black", vertex.label.font=1,vertex.label.dist=0.5,layout=layout_with_fr(Q2.network)) 


```


```{r,echo=FALSE,warning=FALSE,message=FALSE}
edgebundle(Q2.network, padding = 120, fontsize = 10)

```


## Convergence vs. divergence of themes over time

Once again, we binned scores in the the same four groups as above: June, July, Sept/early Oct, and second part of Oct-early Dec.\


```{r,echo=FALSE,warning=FALSE,message=FALSE}

Q2.data.M<-all.data %>% group_by(M)%>%
  summarise(across(social.connectedness_q2:personal.resilience_q2, sum,na.rm=T))


colnames(Q2.data.M)[colnames(Q2.data.M)      # Rename two variable names
                     %in% Q2] <- Q2.q

Q2.data.long.M <- Q2.data.M %>% 
  pivot_longer(-M,names_to = "Theme", values_to = "Count")
Q2.summ.M<-cbind(Q2.data.long.M, Q2.type)
colnames(Q2.summ.M) <- c("Period","Theme","Count","Type")  # change column names
Q2.summ.M <- Q2.summ.M[order(Q2.summ.M$Count), ]  # sort

Q2.data.T<-Q2.summ.M %>% group_by(Period, Type)%>%
    summarise(score=sum(Count))

countByMonth<-all.data %>% group_by(M)%>% 
  summarise(count = n() )

colnames(countByMonth) <- c("Period","N")  # change column names

Q2.data.TT<-full_join(Q2.data.T,countByMonth, by = "Period")
Q2.data.TT$Percent<-(Q2.data.TT$score/Q2.data.TT$N)*100

ggplot(Q2.data.TT, aes(x=Period, y=Percent, group = Type, color=Type, fill=Type)) + 
    geom_line( ) +
    geom_point(shape=21,  size=4) +
   theme_pubclean()+scale_color_d3()+scale_fill_d3()

#BY EACH THEME
#Q2.data.T.ALL<-all.data %>% group_by(M)%>%
#   summarise_at(Q2, funs(sum))
#colnames(Q2.data.T.ALL)[colnames(Q2.data.T.ALL)      # Rename variable names
#                  %in% Q2] <- Q2.q
#
#Q2.data.T.ALL.long <-Q2.data.T.ALL %>%
#    gather(domain, frequency, -c(M), convert = TRUE)
#colnames(Q2.data.T.ALL.long) <- c("Period","Theme","Count")  # change column names
#Q2.data.TT.ALL.long<-full_join(Q2.data.T.ALL.long,countByMonth,, by = "Period")
#Q2.data.TT.ALL.long$Percent<-(Q2.data.TT.ALL.long$Count/Q2.data.TT.ALL.long$N)*100
#ggplot(Q2.data.TT.ALL.long, aes(x=Period, y=Percent, group = Theme, color=Theme, fill=Theme)) + 
#    geom_line( ) +
#  geom_point(shape=21,  size=4) +
#  theme_pubclean()

```

Focus on meta-cognition consistently remains the most frequent recommendation over time (sharing it with prosocial category in September).\
Whereas discussion of societal strategies for wisdom decline in prevalence over time, one strategy clearly picks up in the fall - prosocial strategies!\

One can post-hoc speculate about it. It may be due to greater fatigue and U.S. election, with topics like solidarity becoming more important. It may also have something to do with the fact that prosocial themes equally picked up in answer to the question about type of positive change. \
In other words, scientists forecast greater prosociality OR they recommend greater prosociality. Both communicate some hope for what may be missing in a current society they live in.\


# Wisdom against Negative Consequences

## Summary


We identified 23 distinct categories. 
Except for five themes, categories were mentioned by less than 10% of the interviewees.\
Only two theme was mentioned by at least ten people: Long-term orientation (16%) and Willingness to introduce political-structural change (24%).\

Overall, necessity for political/structure change, long-term focus, and sympathy/compassion are key strategies for mitigating negative consequences of the pandemic, followed by two other social strategies: solidarity, and social support. \
Importantly, ten out of 23 identified themes were meta-cognitive in nature, consistent with recent research on meta-cognition being at the heart of wisdom. It appears leading scholars share this intuition for the context of post-pandemic challenges.\

## Frequency Chart

```{r,echo=FALSE,warning=FALSE,message=FALSE}

Q4 <- c("social.awareness_q4" , "balance.of.personal...others.interest_q4" ,  "solidarity_q4",                    "context.sensitivity_q4",                          "bipartisanship.and.international.cooperation_q4", "critical.thinking_q4",                            "evidence.based.decision.making_q4",               "gratitude_q4",                                   
        "optimism...positivity_q4",                              "improve.communication_q4",                        "long.term.orientation_q4",    "social.support_q4"  ,                  "social.connectedness_q4",               "acknowledge.uncertainty.flexibility_q4",         
        "living.in.the.moment_q4",                         "patience_q4",                                     "perspective.taking_q4",                           "self.reflection.on.what.s.important_q4",          "shared.humanity_q4",                             
        "strive.for.socio.economic.equality_q4",           "sympathy...compassion_q4",                        "political.engagement...structural.change_q4",     "self.distancing_q4") 

Q4.q <- c("Social Awareness" ,"Balance Diverse Interests" ,   "Solidarity"    ,                    
          "Context Sensitivity",  "Polit.Cooperation", "Critical Thinking",  
          "Evidence-based Judgement", "Gratitude","Optimism/Positivity","Improved Communication",                        "Long-term Orientation",   "Social Support","Social Connectedness",
          "Acknowledge Uncertainty",  "Live in the Moment", "Patience",                                     "Perspective-taking", "What's Important?", "Shared Humanity",                             
          "Socio-econ Equality",           "Sympathy&Compassion",                        "Polit.Structural Change",     "Self-distancing") 

Q4.type <- c("Meta-cognition" , "Meta-cognition" ,       "Prosocial"    ,                    
             "Meta-cognition", "Societal", "Meta-cognition","Quest for Knowledge",             
             "Psych.\nWellbeing","Psych.\nWellbeing","Interpersonal","Meta-cognition",
             "Prosocial","Prosocial", "Meta-cognition",   "Psych.\nWellbeing","Self-regulation",                                     "Meta-cognition","Insight","Meta-cognition", "Prosocial",           "Meta-cognition", "Societal",     "Meta-cognition") 

Q4.data<-all.data %>% 
  summarise(across(social.awareness_q4:self.distancing_q4, mean, na.rm=T))

colnames(Q4.data)[colnames(Q4.data)      # Rename two variable names
                     %in% Q4] <- Q4.q

Q4.data.long <- Q4.data %>% 
  pivot_longer(everything(),names_to = "Theme", values_to = "Proportion")
Q4.summ<-cbind(Q4.data.long, Q4.type)
colnames(Q4.summ) <- c("Theme","Proportion","Type")  # change column names
Q4.summ <- Q4.summ[order(Q4.summ$Proportion), ]  # sort
Q4.summ$Theme <- factor(Q4.summ$Theme, levels = Q4.summ$Theme)  # to retain the order in plot
Q4.summ$Percent<-Q4.summ$Proportion*100
Q4.plot<-ggplot(Q4.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() +
  geom_bar(stat="identity", width=.5) + ggtitle("Wisdom against Negative Consequences")+theme_pubclean()+scale_fill_d3()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom", legend.text=element_text(size=6.5))+labs(x="",y="%")
Q4.plot
```

## How many themes per person?

Did people just report 1-2 themes or a great number of themes? How do such trends vary across people?\
It turns out most people mentioned just one or two (median) negative consequences. Two people mentioned up to 5 and 6 themes, respectively.\


```{r,echo=FALSE,warning=FALSE,message=FALSE}
all.data$Q4.pp<-rowSums(all.data[Q4])
ggplot(all.data, aes(x = Q4.pp)) +   geom_vline(xintercept=median(all.data$Q4.pp, na.rm=T), color="red")+
  geom_bar()+theme_pubclean()+scale_fill_d3()+labs(x="Number of Themes per Person",y="Count")
           
```

## Frequencies by Gender and Field

Once again, it is not surprising that wisdom experts mention meta-cognitive and moral considerations. The question is whether scientists without much familiarity with the science of wisdom would mention similar constructs for this question, too. It appears that is the case, with non-wisdom experts emphasizing such meta-cognitive categories as long-term orientation, sympathy and compassion, context sensitivity, appreciation of the idea of shared humanity, critical thinking and acknowledging uncertainty, along with moral aspirations about solidarity, social support, and paying greater attention to one's family and relationships, and promoting political cooperation and societal change toward fair and just society. In fact, the latter categories were strongly prevalent among non-wisdom experts. 


```{r,echo=FALSE,warning=FALSE,message=FALSE}
##BY gender
Q4.data.G<-all.data %>% group_by(Gender)%>% 
  summarise(across(social.awareness_q4:self.distancing_q4, mean, na.rm=T))

colnames(Q4.data.G)[colnames(Q4.data.G)      # Rename two variable names
                    %in% Q4] <- Q4.q

Q4.data.G.long <- Q4.data.G %>% 
  pivot_longer(-Gender,names_to = "Theme", values_to = "Proportion")
Q4.G.summ<-cbind(Q4.data.G.long, Q4.type)
colnames(Q4.G.summ) <- c("Gender","Theme","Proportion","Type")  # change column names
Q4.G.summ$Percent<-Q4.G.summ$Proportion*100
Q4.G.plot<-ggplot(Q4.G.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Gender)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Wisdom against Consequences")+theme_pubclean()+scale_fill_d3()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom", legend.text=element_text(size=6.5))+labs(x="",y="%")
Q4.G.plot


##BY wisdom familiarity
Q4.data.W<-all.data %>% group_by(FamiliarWisdom)%>% 
  summarise(across(social.awareness_q4:self.distancing_q4, mean, na.rm = TRUE))

colnames(Q4.data.W)[colnames(Q4.data.W)      # Rename two variable names
                     %in% Q4] <- Q4.q

Q4.data.W.long <- Q4.data.W %>% 
  pivot_longer(-FamiliarWisdom,names_to = "Theme", values_to = "Proportion")
Q4.W.summ<-cbind(Q4.data.W.long, Q4.type)
colnames(Q4.W.summ) <- c("FamiliarWisdom","Theme","Proportion","Type")  # change column names
Q4.W.summ$Percent<-Q4.W.summ$Proportion*100
Q4.W.plot<-ggplot(Q4.W.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ FamiliarWisdom)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Wisdom against Negative Consequences\nby Wisdom Familarity")+theme_pubclean()+scale_fill_d3()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom", legend.text=element_text(size=6.5))+labs(x="",y="%")
Q4.W.plot

#test whether familiarity with wisdom impacts type of themes selected
Q4.data.Famil.long<-all.data %>% pivot_longer(cols=social.awareness_q4:self.distancing_q4, names_to="Theme",values_to="Presence")

Q4.Famil.result<-glmer(Presence ~ Theme*FamiliarWisdom + (1 | Name), data = Q4.data.Famil.long, family = binomial)
summary(Q4.Famil.result)
Anova(Q4.Famil.result,type="III")
#no significant difference by familiarity with wisdom scholarship

##BY socio-cultural vs. non-socio-cultural expertise
Q4.data.SC<-all.data %>% group_by(Field)%>% 
  summarise(across(social.awareness_q4:self.distancing_q4, mean, na.rm=T))

colnames(Q4.data.SC)[colnames(Q4.data.SC)      # Rename two variable names
                     %in% Q4] <- Q4.q

Q4.data.SC.long <- Q4.data.SC %>% 
  pivot_longer(-Field,names_to = "Theme", values_to = "Proportion")
Q4.SC.summ<-cbind(Q4.data.SC.long, Q4.type)
colnames(Q4.SC.summ) <- c("Field","Theme","Proportion","Type")  # change column names
Q4.SC.summ$Percent<-Q4.SC.summ$Proportion*100
Q4.SC.plot<-ggplot(Q4.SC.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Field)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Wisdom against Negative Consequences")+theme_pubclean()+scale_fill_d3()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom", legend.text=element_text(size=6.5))+labs(x="",y="%")
Q4.SC.plot

Q4.SC.result<-glmer(Presence ~ Theme*Field + (1 | Name), data = Q4.data.Famil.long, family = binomial)
summary(Q4.SC.result)
Anova(Q4.SC.result,type="III")


```


## Frequencies by Cognitive Style: Dialectical Thinking

### Is there a difference in themes mentioned between experts who show dialecticism in their reasoning about societal change vs. those who do not invoke dialectism?

It appears that **people invoking dialecticism** in their reasoning were more likely to raise the themes of *perspective-taking*, *self-distancing*, *focus on the concept of shared humanity*,*solidarity*, and *optimism/positivity*  compared to those who did not show dialecticism.\
\

```{r,echo=FALSE,warning=FALSE,message=FALSE}
#Dialectical thinking

Q4.data.Dia<-all.data %>% group_by(Dialectic_Final)%>% 
  summarise(across(social.awareness_q4:self.distancing_q4, mean, na.rm=TRUE))

colnames(Q4.data.Dia)[colnames(Q4.data.Dia)      # Rename two variable names
                     %in% Q4] <- Q4.q
Q4.data.Dia$Dialecticism[Q4.data.Dia$Dialectic_Final==0]<-"Not Dialectical"
Q4.data.Dia$Dialecticism[Q4.data.Dia$Dialectic_Final==1]<-"Dialectical"

Q4.data.Dia.long <- Q4.data.Dia %>% 
  pivot_longer(-c("Dialectic_Final","Dialecticism"),names_to = "Theme", values_to = "Proportion")
Q4.Dia.summ<-cbind(Q4.data.Dia.long, Q4.type)
colnames(Q4.Dia.summ) <- c("Dialectical Code","Dialecticism","Theme","Proportion","Type")  # change column names
Q4.Dia.summ$Percent<-Q4.Dia.summ$Proportion*100
Q4.Dia.plot<-ggplot(Q4.Dia.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Dialecticism)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Wisdom for Negative Consequences by Cognitive Style")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="%")
Q4.Dia.plot


```


### Is there a difference in **Common Wisdom Model** themes between dialectical experts vs. those who do not invoke dialectism?

No group difference. We observe a main effect: both groups are sig more likely to favor CWM vs. other themes.


```{r,warning=FALSE,message=FALSE}
#Are dialectical thinkers more likely to mention core aspects of wisdom as identified by the Common wisdom model per Toronto Wisdom Task Force 2019?

#sum up all morals
moralsQ4<-c( "solidarity_q4",                        "bipartisanship.and.international.cooperation_q4",                  "gratitude_q4",  "social.support_q4"  ,                  "social.connectedness_q4",             
             "shared.humanity_q4",      "strive.for.socio.economic.equality_q4",               "political.engagement...structural.change_q4") 


#sum up all metacognitive
metacogQ4<- c("social.awareness_q4" , "balance.of.personal...others.interest_q4" ,  "context.sensitivity_q4",     "critical.thinking_q4",       "long.term.orientation_q4",             "acknowledge.uncertainty.flexibility_q4",         
               "perspective.taking_q4",                          "sympathy...compassion_q4",                         "self.distancing_q4") 
 
#sum the rest
othersQ4<-c("evidence.based.decision.making_q4",    "optimism...positivity_q4",                              "improve.communication_q4",  "living.in.the.moment_q4",                         "patience_q4",    "self.reflection.on.what.s.important_q4") 

all.data$morals.Q4<-rowSums(all.data[moralsQ4])
all.data$metacog.Q4<-rowSums(all.data[metacogQ4])
all.data$others.Q4<-rowSums(all.data[othersQ4])

all.data$CWM.Q4<-rowSums(all.data[c("morals.Q4","metacog.Q4")])

Q4.dia.data<-all.data %>% pivot_longer(cols=morals.Q4:others.Q4, names_to="type",values_to="count")

Q4.dia.result<-glmer(count ~ type*Dialecticism + (1 | Name), data = Q4.dia.data, family = poisson)
Anova(Q4.dia.result,type="III") #no significant differences
sjPlot::plot_model(Q4.dia.result,type="int")

Q4.dia.result.emt<-emmeans::emmeans(Q4.dia.result,"type",var="Dialecticism")
pairs(Q4.dia.result.emt)

#contrast of cWM (morals and meta-cog) vs. the rest

Q4.dia.data.C<-all.data %>% pivot_longer(cols=c("others.Q4","CWM.Q4"), names_to="type",values_to="count")

Q4.dia.result.C<-glmer(count ~ type*Dialecticism + (1 | Name), data = Q4.dia.data.C, family = poisson)
Anova(Q4.dia.result.C,type="III") #no significant differences
sjPlot::plot_model(Q4.dia.result.C,type="int")

Q4.dia.result.C.emt<-emmeans::emmeans(Q4.dia.result.C,"type",by="Dialecticism")
pairs(Q4.dia.result.C.emt)




Q4.data.Dia.CWM<-all.data %>% group_by(Dialecticism)%>% 
  summarise(across(morals.Q4:others.Q4, sum, na.rm=TRUE))%>%
pivot_longer(-Dialecticism,names_to = "Theme", values_to = "Count")

Q4.Dia.CWM.plot<- ggplot(Q4.data.Dia.CWM,aes(x=Dialecticism,y=Count,fill=Theme))+
  geom_col(stat="identity")+theme_pubclean()+scale_fill_d3()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="Count")

Q4.Dia.CWM.plot #more meta-cognition and more morals among nondialectics (by virtue of them bringing up more themes in general and being a larger sample)
 

mosaic(xtabs(Count~ Dialecticism+Theme, data= Q4.data.Dia.CWM), shade=TRUE) #but not significant


Q4.data.W.CWM<-all.data %>% group_by(FamiliarWisdom)%>% 
  summarise(across(moralsQ4:othersQ4, sum, na.rm=TRUE))%>%
pivot_longer(-FamiliarWisdom,names_to = "Theme", values_to = "Count")
mosaic(xtabs(Count~  FamiliarWisdom+Theme, data= Q4.data.W.CWM), shade=TRUE) #those familiar with wisdom research are significantly more likely to mention meta-cognition than those not familiar with wisdom research.

chisq.test(xtabs(Count~  FamiliarWisdom+Theme, data= Q4.data.W.CWM))
Q4.famil.result<-glmer(count ~ type*FamiliarWisdom + (1 | Name), data = Q4.dia.data, family = poisson)
Anova(Q4.famil.result,type="III") #no significant differences
sjPlot::plot_model(Q4.famil.result,type="int")

Q4.famil.result.emt<-emmeans::emmeans(Q4.famil.result,"type",by="FamiliarWisdom")
pairs(Q4.famil.result.emt) #  folks familiar with wisdom research more likely to mention metacog vs. other categories, whereas folks not familiar with wisdom more likely to mention morals vs. others. No sig differences in preference for meta-cog vs. morals in both groups.


```

## PCA 

```{r,echo=FALSE,warning=FALSE,message=FALSE}

vss(all.data[Q4], rotate="oblimin", fm="pc")
principal(all.data[Q4], nfactors = 8, rotate="oblimin")


```

Similar to prior results, we can address the question of diversity by examining the degree to which scores across themes are reducible to common component, examining principal component analysis (PCA).\
It appears that 8 components is as good a start to account for 23 themes for Question 4 (wisdom against neg consequences), as any other. Once again, the scree plot line is pretty flat.\
Again, keep in mind that reducing 23 items to 8 components is not parsimonious. Moreover, when we reduce the items to 8 components, the first component explains only 10 % of the variance, and in total 8 components explain 60% of variance.\
Given that each theme by default explains 4.4% of the variance (1/23), this is not much. \
Furthermore, each of these components is largely based on 1-2 items (loadings > .6). \

In short, principal components show substantial diversity, with the top factor political engagement and structural change not even loading squarely on any of the components (and negatively loading on the component it has highest loading for - the one with positive loading for self-distancing).\


## Network Graph

Even after removing negligible correlations (r < .17), cluster analyses on top of the network graphs show four clusters, three of which are strongly inter-related (perspective-taking, balance of diverse interests, and socio-econ equality): \
•	Four clusters: \
•	Context-sensitivity & perspective-taking \
•	Meta-cognition : balance, self-distancing, focus on the big picture shared humanity \
•	Political engagement and structural change, cooperation, as well as fight against inequality, linked with patience /long-term focus \
•	Psych well-being enhancement: Gratitude, acknowledge uncertainty and focus on the present \

Balance of diverse interests appears in the center of everything, connecting to other themes mentioned by experts.\

```{r,echo=FALSE,warning=FALSE,message=FALSE}
Q4.cor<-partial.r(all.data,Q4,cs(tokens.Q4),method="spearman")

row.names(Q4.cor) <-Q4.q
colnames(Q4.cor) <-Q4.q

# Keep only meaningful correlations (<.17)
Q4.cor[Q4.cor<0.17] <- 0

# Make an Igraph object from this matrix:
Q4.network <- graph_from_adjacency_matrix( Q4.cor, weighted=T, mode="undirected", diag=F)



#color palette
coul <- brewer.pal(nlevels(as.factor(Q4.type)), "Dark2")

# Map the color to types
my_color <- coul[as.numeric(as.factor(Q4.type))]

# plot with edges

plot(Q4.network,
     edge.color=rep(c("red","pink"),5),           # Edge color
     edge.width=seq(1,10),                        # Edge width, defaults to 1
     edge.arrow.size=1,                           # Arrow size, defaults to 1
     edge.arrow.width=1,                          # Arrow width, defaults to 1
     edge.lty=c("solid"),
     vertex.label.color="black",
     vertex.color=my_color
     # Line type, could be 0 or "blank", 1 or "solid", 2 or "dashed", 3 or "dotted", 4 or "dotdash", 5 or "longdash", 6 or "twodash"
     #edge.curved=c(rep(0,5), rep(1,5))            # Edge curvature, range 0-1 (FALSE sets it to 0, TRUE to 0.5)
)

# title and legend
#legend("topleft", 
#       legend=paste(levels(as.factor(Q4.type))), 
#       col = coul , 
#       bty = "n", pch=20 , pt.cex = 2, cex = 1,
#       text.col="black" , horiz = F)

##KEY NETWORK PLOT#####


Q4.ceb <- cluster_edge_betweenness(Q4.network) 

dendPlot(Q4.ceb, mode="hclust")
edge.start <- ends(Q4.network, es=E(Q4.network), names=F)[,1]
edge.col <- V(Q4.network)$color[edge.start]

plot(Q4.ceb, Q4.network, axes=F, ,vertex.shape="none",edge.color=edge.col, edge.curved=.1,
     vertex.label.color="black", vertex.label.font=1,vertex.label.dist=0.5,layout=layout_with_fr(Q4.network)) 


```



```{r,echo=FALSE,warning=FALSE,message=FALSE}
edgebundle(Q4.network, padding = 120, fontsize = 10)

```


## Convergence vs. divergence of themes over time


```{r,echo=FALSE,warning=FALSE,message=FALSE}

Q4.data.M<-all.data %>% group_by(M)%>%
  summarise(across(social.awareness_q4:self.distancing_q4, sum,na.rm=T))


colnames(Q4.data.M)[colnames(Q4.data.M)      # Rename two variable names
                     %in% Q4] <- Q4.q

Q4.data.long.M <- Q4.data.M %>% 
  pivot_longer(-M,names_to = "Theme", values_to = "Count")
Q4.summ.M<-cbind(Q4.data.long.M, Q4.type)
colnames(Q4.summ.M) <- c("Period","Theme","Count","Type")  # change column names
Q4.summ.M <- Q4.summ.M[order(Q4.summ.M$Count), ]  # sort

Q4.data.T<-Q4.summ.M %>% group_by(Period, Type)%>%
    summarise(score=sum(Count))

countByMonth<-all.data %>% group_by(M)%>% 
  summarise(count = n() )

colnames(countByMonth) <- c("Period","N")  # change column names

Q4.data.TT<-full_join(Q4.data.T,countByMonth, by = "Period")
Q4.data.TT$Percent<-(Q4.data.TT$score/Q4.data.TT$N)*100

ggplot(Q4.data.TT, aes(x=Period, y=Percent, group = Type, color=Type, fill=Type)) + 
    geom_line( ) +
    geom_point(shape=21,  size=4) +
   theme_pubclean()+scale_color_d3()+scale_fill_d3()

#BY EACH THEME
#Q4.data.T.ALL<-all.data %>% group_by(M)%>%
#   summarise_at(Q4, funs(sum))
#colnames(Q4.data.T.ALL)[colnames(Q4.data.T.ALL)      # Rename variable names
#                  %in% Q4] <- Q4.q
#
#Q4.data.T.ALL.long <-Q4.data.T.ALL %>%
#    gather(domain, frequency, -c(M), convert = TRUE)
#colnames(Q4.data.T.ALL.long) <- c("Period","Theme","Count")  # change column names
#Q4.data.TT.ALL.long<-full_join(Q4.data.T.ALL.long,countByMonth,, by = "Period")
#Q4.data.TT.ALL.long$Percent<-(Q4.data.TT.ALL.long$Count/Q4.data.TT.ALL.long$N)*100
#ggplot(Q4.data.TT.ALL.long, aes(x=Period, y=Percent, group = Theme, color=Theme, fill=Theme)) + 
#    geom_line( ) +
#  geom_point(shape=21,  size=4) +
#  theme_pubclean()

```

Once again, focus on meta-cognition consistently remains the most frequent recommendation over time and in fact, increases in importance over time.\
As with wisdom to sustain positive change (Q2), discussion of societal strategies for wisdom against negative changes is less prevalent over time, whereas prosocial tendencies becoming relatively more prevant over time!\

One can post-hoc speculate about it. It may be due to greater fatigue and U.S. election, with topics like solidarity becoming more important and focus becoming increasingly within-country centric and less about international cooperation (which would be captured by societal theme). This said both societal and prosocial themes tap into the same general foundation of wisdom - moral aspirations (Grossmannn et al., 2020) - it captures both prosocial concerns on the interpersonal level and broader societal concerns with bipartisanship and the common good. \

# Wisdom Now


## Summary


We identified 24 distinct categories.\
Except for four themes, categories were mentioned by less than 10% of the interviewees.\
Only two theme was mentioned by at least ten people: Agency & Control (18%) and Social connectedness (25%).\

Once again, seven categories were meta-cognitive, whereas 3 were (pro)social and one was societal.\



## Frequency Chart

```{r,echo=FALSE,fig.width = 10,fig}
###Q5################################################


Q5 <- c("balance.of.personal...others.interest_q5",       
 "clear.governmental.communication_q5",             "solidarity_q5",                                   "bipartisanship.and.international.cooperation_q5",
 "critical.thinking_q5",                            "embrace.new.tech_q5",                             "agency...control_q5",                            
 "follow.rules_q5",                                 "optimism...positivity_q5",                        "learning.from.the.pandemics_q5",                  "improve.communication_q5",
 "prosocial.behavior_q5",                           "long.term.orientation_q5",         "acknowledge.uncertainty.flexibility_q5",          "living.in.the.moment_q5",
 "patience_q5",                                     "perspective.taking_q5",                           "self.reflection.on.what.is.important_q5",         "social.connectedness_q5",                        
 "self.distancing_q5",                              "strive.for.socio.economic.equality_q5",           "sympathy...compassion_q5",                        "take.science.seriously_q5",                       "personal.resilience_q5") 




Q5.q <- c("Balance Diverse Interests",
          "Clear Government\nCommunication"  ,           "Solidarity"      ,                   "Polit.Cooperation",
        "Critical Thinking" ,                           "Embrace New Tech"   , "Agency/Control" ,                            
        "Follow Rules",                                 "Optimism/Positivity"  ,         "Learning from Pandemics" ,    "Improved Communication",
          "Prosocial Behavior"   ,                        "Long-term Orientation"     ,                   "Acknowledge Uncertainty"  ,        "Live in the Moment" ,                       
        "Patience"        ,                             "Perspective-taking" ,                          "What's Important?" , "Social Connectedness",       
        "Self-distancing" ,   "Socio-econ Equality"     ,      "Sympathy&Compassion"          ,              "Take Science Seriously" ,   "Personal Resilience")




Q5.type <- c("Meta-cognition",  "Quest\nfor Knowledge"  ,           "Prosocial"      ,                   "Societal",
             "Meta-cognition" ,                           "Work"   ,                         
             "Agency/Control" ,                            "Agency/Control",                                 "Psych.\nWellbeing"  ,                            "Quest for Knowledge" ,                 "Interpersonal", 
             
             "Prosocial"   ,                        "Meta-cognition"     ,                   "Meta-cognition"  ,        "Psych.\nWellbeing" ,                       
             "Self-regulation"        ,                             "Meta-cognition" ,                          "Insight"  , "Prosocial"     ,        "Meta-cognition" ,                            
             "Prosocial"     ,      "Meta-cognition"          ,              "Quest\nfor Knowledge" , "Psych.\nWellbeing" )


Q5.data<-all.data %>% 
  summarise(across(balance.of.personal...others.interest_q5:personal.resilience_q5, sum, na.rm=T))


colnames(Q5.data)[colnames(Q5.data)      # Rename two variable names
                     %in% Q5] <- Q5.q

Q5.data.long <- Q5.data %>% 
  pivot_longer(everything(),names_to = "Theme", values_to = "Count")
Q5.summ<-cbind(Q5.data.long, Q5.type)
colnames(Q5.summ) <- c("Theme","Count","Type")  # change column names
Q5.summ <- Q5.summ[order(Q5.summ$Count), ]  # sort
Q5.summ$Theme <- factor(Q5.summ$Theme, levels = Q5.summ$Theme)  # to retain the order in plot
Q5.summ$Percent<-Q5.summ$Count/length(all.data$Q5)*100
Q5.plot<-ggplot(Q5.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() +
  geom_bar(stat="identity", width=.5) + ggtitle("Wisdom Now")+theme_pubclean()+scale_fill_manual(values = brewer.pal(n = 11, name = "Set3"))+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom")+labs(x="",y="%")
Q5.plot


#create the graph for Conversation article

ggplot(Q5.summ, aes(x=Theme, y=Percent, color=Type)) +
  geom_segment(aes(x=Theme, xend=Theme, y=0, yend=Percent), color="black") +
  geom_point( size=4, alpha=0.9) +
  theme_light() +
  coord_flip() +scale_color_d3(palette = "category20c")+
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
   legend.position="none")
  


```






## How many themes per person?

Did people just report 1-2 themes or a great number of themes? How do such trends vary across people?\

Most people reported 1-2 themes, with the max being 4 themes mentioned by 4 people.

```{r,echo=FALSE,warning=FALSE,message=FALSE}
all.data$Q5.pp<-rowSums(all.data[Q5])

ggplot(all.data, aes(x = Q5.pp)) +   geom_vline(xintercept=median(all.data$Q5.pp,na.rm=T), color="red")+
  geom_bar()+theme_pubclean()+scale_fill_d3()+labs(x="Number of Themes per Person",y="Count")
           
```


The majority of people mention only one or two themes here. The median was two themes (as indicated by vertical red line). One-fifth mentioned three themes, and only 3 people mentioned 4 themes (and one person mentioned five themes). 


## Frequencies by Gender and Field

Once again, it is not surprising that wisdom experts mention meta-cognitive and moral considerations. The question is whether scientists without much familiarity with the science of wisdom would mention similar constructs for this question, too. It appears that is the case, with non-wisdom experts emphasizing meta-cognitive categories such as sympathy and compassion, long-term orientation, context sensitivity, critical thinking and acknowledging uncertainty, along with moral aspirations about solidarity, and paying greater attention to one's family and relationship. In fact, the latter categories were strongly prevalent among non-wisdom experts. 

```{r,echo=FALSE,warning=FALSE,message=FALSE}
##BY gender
Q5.data.G<-all.data %>% group_by(Gender)%>% 
  summarise(across(balance.of.personal...others.interest_q5:personal.resilience_q5, mean, na.rm=T))

colnames(Q5.data.G)[colnames(Q5.data.G)      # Rename two variable names
                     %in% Q5] <- Q5.q

Q5.data.G.long <- Q5.data.G %>% 
  pivot_longer(-Gender,names_to = "Theme", values_to = "Proportion")
Q5.G.summ<-cbind(Q5.data.G.long, Q5.type)
colnames(Q5.G.summ) <- c("Gender","Theme","Proportion","Type")  # change column names
Q5.G.summ$Percent<-Q5.G.summ$Proportion*100
Q5.G.plot<-ggplot(Q5.G.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Gender)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Wisdom Now")+theme_pubclean()+scale_fill_manual(values = brewer.pal(n = 11, name = "Set3"))+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom", legend.text=element_text(size=7))+labs(x="",y="%")
Q5.G.plot



##BY familiarity with wisdom
Q5.data.W<-all.data %>% group_by(FamiliarWisdom)%>% 
  summarise(across(balance.of.personal...others.interest_q5:personal.resilience_q5, mean, na.rm=T))

colnames(Q5.data.W)[colnames(Q5.data.W)      # Rename two variable names
                     %in% Q5] <- Q5.q

Q5.data.W.long <- Q5.data.W %>% 
  pivot_longer(-FamiliarWisdom,names_to = "Theme", values_to = "Proportion")
Q5.W.summ<-cbind(Q5.data.W.long, Q5.type)
colnames(Q5.W.summ) <- c("FamiliarWisdom","Theme","Proportion","Type")  # change column names
Q5.W.summ$Percent<-Q5.W.summ$Proportion*100
Q5.W.plot<-ggplot(Q5.W.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ FamiliarWisdom)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Wisdom Now by Wisdom Familiarity")+theme_pubclean()+scale_fill_manual(values = brewer.pal(n = 11, name = "Set3"))+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom", legend.text=element_text(size=7))+labs(x="",y="%")
Q5.W.plot

#test whether familiarity with wisdom impacts type of themes selected
Q5.data.Famil.long<-all.data %>% pivot_longer(cols=balance.of.personal...others.interest_q5:personal.resilience_q5, names_to="Theme",values_to="Presence")

Q5.Famil.result<-glmer(Presence ~ Theme*FamiliarWisdom + (1 | Name), data = Q5.data.Famil.long, family = binomial)
summary(Q5.Famil.result)
Anova(Q5.Famil.result,type="III")
#no significant difference by familiarity with wisdom scholarship

##BY socio-cultural vs. non-socio-cultural expertise
Q5.data.SC<-all.data %>% group_by(Field)%>% 
  summarise(across(balance.of.personal...others.interest_q5:personal.resilience_q5, mean, na.rm=T))

colnames(Q5.data.SC)[colnames(Q5.data.SC)      # Rename two variable names
                     %in% Q5] <- Q5.q

Q5.data.SC.long <- Q5.data.SC %>% 
  pivot_longer(-Field,names_to = "Theme", values_to = "Proportion")
Q5.SC.summ<-cbind(Q5.data.SC.long, Q5.type)
colnames(Q5.SC.summ) <- c("Field","Theme","Proportion","Type")  # change column names
Q5.SC.summ$Percent<-Q5.SC.summ$Proportion*100
Q5.SC.plot<-ggplot(Q5.SC.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Field)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Wisdom Now")+theme_pubclean()+scale_fill_manual(values = brewer.pal(n = 11, name = "Set3"))+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom", legend.text=element_text(size=7))+labs(x="",y="%")
Q5.SC.plot


Q5.SC.result<-glmer(Presence ~ Theme*Field + (1 | Name), data = Q5.data.Famil.long, family = binomial)
summary(Q5.SC.result)
Anova(Q5.SC.result,type="III")


```


## Frequencies by Cognitive Style: Dialectical Thinking

Is there a difference in themes mentioned between experts who show dialecticism in their reasoning about societal change vs. those who do not invoke dialectism?\

It appears that **people invoking dialecticism** in their reasoning were quite similar to those not invoking dialecticism.\
\
There are were a noticeable differences: *dialectics* more likely to raise the themes of *self-distancing*, *balancing diverse interests*, *prosocial behavior*, *clear governmental communication*, and *following rules* compared to those who did not show dialecticism.\
\

```{r,echo=FALSE,warning=FALSE,message=FALSE}
#Dialectical thinking

Q5.data.Dia<-all.data %>% group_by(Dialectic_Final)%>% 
  summarise(across(balance.of.personal...others.interest_q5:personal.resilience_q5, mean, na.rm=TRUE))

colnames(Q5.data.Dia)[colnames(Q5.data.Dia)      # Rename two variable names
                     %in% Q5] <- Q5.q
Q5.data.Dia$Dialecticism[Q5.data.Dia$Dialectic_Final==0]<-"Not Dialectical"
Q5.data.Dia$Dialecticism[Q5.data.Dia$Dialectic_Final==1]<-"Dialectical"

Q5.data.Dia.long <- Q5.data.Dia %>% 
  pivot_longer(-c("Dialectic_Final","Dialecticism"),names_to = "Theme", values_to = "Proportion")
Q5.Dia.summ<-cbind(Q5.data.Dia.long, Q5.type)
colnames(Q5.Dia.summ) <- c("Dialectical Code","Dialecticism","Theme","Proportion","Type")  # change column names
Q5.Dia.summ$Percent<-Q5.Dia.summ$Proportion*100
Q5.Dia.plot<-ggplot(Q5.Dia.summ, aes(x=Theme, y=Percent,fill=Type)) + coord_flip() + facet_grid(. ~ Dialecticism)+
  geom_bar(stat="identity", width=.5,position=position_dodge()) + ggtitle("Wisdom Now by Cognitive Style")+theme_pubclean()+scale_fill_aaas()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="%")
Q5.Dia.plot

```

### Is there a difference in **Common Wisdom Model** themes between dialectical experts vs. those who do not invoke dialectism?

It appears there is a difference in the direction of dialectical experts being likely to mention moral aspirations and meta-cognition, but it is not significant (p = .0597). No significant differences between experts familiar with wisdom research and those not familiar.

```{r,warning=FALSE,message=FALSE}
#Are dialectical thinkers more likely to mention core aspects of wisdom as identified by the Common wisdom model per Toronto Wisdom Task Force 2019?

#sum up all morals
moralsQ5<-c("solidarity_q5",                                   "bipartisanship.and.international.cooperation_q5",
 "follow.rules_q5",                                
 "prosocial.behavior_q5",                           
 "social.connectedness_q5",                         "strive.for.socio.economic.equality_q5",         "take.science.seriously_q5") 


#sum up all metacognitive
metacogQ5<- c("balance.of.personal...others.interest_q5",       
  "critical.thinking_q5",                                              
        "long.term.orientation_q5",         "acknowledge.uncertainty.flexibility_q5",                "perspective.taking_q5",                                         
 "self.distancing_q5",                              "sympathy...compassion_q5"                      ) 
 
#sum the rest
othersQ5<-c("clear.governmental.communication_q5",   "embrace.new.tech_q5",                             "agency...control_q5",                            
        "optimism...positivity_q5",                        "learning.from.the.pandemics_q5",                  "improve.communication_q5",
      "living.in.the.moment_q5", "patience_q5","self.reflection.on.what.is.important_q5",             "personal.resilience_q5") 

all.data$morals.Q5<-rowSums(all.data[moralsQ5])
all.data$metacog.Q5<-rowSums(all.data[metacogQ5])
all.data$others.Q5<-rowSums(all.data[othersQ5])

all.data$CWM.Q5<-rowSums(all.data[c("morals.Q5","metacog.Q5")])

Q5.dia.data<-all.data %>% pivot_longer(cols=morals.Q5:others.Q5, names_to="type",values_to="count")

Q5.dia.result<-glmer(count ~ type*Dialecticism + (1 | Name), data = Q5.dia.data, family = poisson)
Anova(Q5.dia.result,type="III") #no significant differences
sjPlot::plot_model(Q5.dia.result,type="int")

Q5.dia.result.emt<-emmeans::emmeans(Q5.dia.result,"type",var="Dialecticism")
pairs(Q5.dia.result.emt)

#contrast of cWM (morals and meta-cog) vs. the rest

Q5.dia.data.C<-all.data %>% pivot_longer(cols=c("others.Q5","CWM.Q5"), names_to="type",values_to="count")

Q5.dia.result.C<-glmer(count ~ type*Dialecticism + (1 | Name), data = Q5.dia.data.C, family = poisson)
Anova(Q5.dia.result.C,type="III") #no significant differences
sjPlot::plot_model(Q5.dia.result.C,type="int")

Q5.dia.result.C.emt<-emmeans::emmeans(Q5.dia.result.C,"type",by="Dialecticism")
pairs(Q5.dia.result.C.emt)


Q5.data.Dia.CWM<-all.data %>% group_by(Dialecticism)%>% 
  summarise(across(morals.Q5:others.Q5, sum, na.rm=TRUE))%>%
pivot_longer(-Dialecticism,names_to = "Theme", values_to = "Count")

Q5.Dia.CWM.plot<- ggplot(Q5.data.Dia.CWM,aes(x=Dialecticism,y=Count,fill=Theme))+
  geom_col(stat="identity")+theme_pubclean()+scale_fill_d3()+theme(plot.title = element_text(hjust = 0.5),legend.position="bottom",legend.text=element_text(size=8))+labs(x="",y="Count")

Q5.Dia.CWM.plot #more meta-cognition and more morals among dialectics
 

mosaic(xtabs(Count~  Dialecticism+Theme, data= Q5.data.Dia.CWM), shade=TRUE) #but not significant


Q5.data.W.CWM<-all.data %>% group_by(FamiliarWisdom)%>% 
  summarise(across(moralsQ5:othersQ5, sum, na.rm=TRUE))%>%
pivot_longer(-FamiliarWisdom,names_to = "Theme", values_to = "Count")

mosaic(xtabs(Count~ FamiliarWisdom+Theme, data= Q5.data.W.CWM), shade=TRUE) #those familiar with wisdom research are more likely to mention meta-cognition, but not significantly so.

chisq.test(xtabs(Count~ FamiliarWisdom+Theme, data= Q5.data.W.CWM))

Q5.famil.result<-glmer(count ~ type*FamiliarWisdom + (1 | Name), data = Q5.dia.data, family = poisson)
Anova(Q5.famil.result,type="III") #no significant differences
sjPlot::plot_model(Q5.famil.result,type="int")

Q5.famil.result.emt<-emmeans::emmeans(Q5.famil.result,"type",by="FamiliarWisdom")
pairs(Q5.famil.result.emt)

pairs(emmeans::emmeans(Q5.famil.result,"FamiliarWisdom",by="type"))

```


## PCA 

```{r,echo=FALSE,warning=FALSE,message=FALSE}

vss(all.data[Q5], rotate="oblimin", fm="pc")
principal(all.data[Q5], nfactors = 8, rotate="oblimin")


```

Similar to prior results, we can address the question of diversity by examining the degree to which scores across themes are reducible to common component, examining principal component analysis (PCA).\
It appears that 8 components is as good a start to account for 25 themes for Question 5 (wisdom now), as any other. Once again, the scree plot line is pretty flat.\
Again, keep in mind that reducing 24 items to 8 components is not parsimonious. Moreover, when we reduce the items to 8 components, the first component explains only 9 % of the variance, and in total 8 components explain 57% of variance.\
Given that each theme by default explains 4.2% of the variance (1/23), this is not much. \
Furthermore, each of these components is largely based on 1-2 items (loadings > .6), with exception of TC2 : bipartisanship & int.cooperation, learning from the pandemic, long-term orientation.\

In short, principal components show substantial diversity, with the top factor social connectedness not even loading squarely on any of the components (and negatively loading on the component it has highest loading for - the one with positive loading for critical thinking).\


## Network Graph

Even after removing negligible correlations (r < .17), cluster analyses on top of the network graphs show five clusters, four of which are strongly inter-related (embrace new tech. and optimism/positivity, solidarity and prosocial behavior, self-distancing and pol. cooperation): \

Five clusters: \
*	Solidarity & pol. cooperation\
*	Resilience-building: Establish agency, patience, prosociality,, optimism \
* Focus on here & now (incl. role of new tech.)\
*	Meta-cognition: self-distancing, balancing different interests, critical thinking\
*	Social capital: social connectedness, improved communication, socio-econ equality, sympathy & compassion\


```{r,echo=FALSE,warning=FALSE,message=FALSE}
#create new variables, adjusting for number of words in each narrative
Q5.cor<-partial.r(all.data,Q5,cs(tokens.Q5),method="spearman")

row.names(Q5.cor) <-Q5.q
colnames(Q5.cor) <-Q5.q

# Keep only meaningful correlations (<.17)
Q5.cor[is.na(Q5.cor)==TRUE] <- 0
Q5.cor[Q5.cor<0.17] <- 0

# Make an Igraph object from this matrix:
Q5.network <- graph_from_adjacency_matrix( Q5.cor, weighted=T, mode="undirected", diag=F)


#color palette
coul <- brewer.pal(nlevels(as.factor(Q5.type)), "Set3")

# Map the color to types
my_color <- coul[as.numeric(as.factor(Q5.type))]

# plot with edges

plot(Q5.network,
     edge.color=rep(c("red","pink"),5),           # Edge color
     edge.width=seq(1,10),                        # Edge width, defaults to 1
     edge.arrow.size=1,                           # Arrow size, defaults to 1
     edge.arrow.width=1,                          # Arrow width, defaults to 1
     edge.lty=c("solid"),
     vertex.label.color="black",
     vertex.color=my_color
     # Line type, could be 0 or "blank", 1 or "solid", 2 or "dashed", 3 or "dotted", 4 or "dotdash", 5 or "longdash", 6 or "twodash"
     #edge.curved=c(rep(0,5), rep(1,5))            # Edge curvature, range 0-1 (FALSE sets it to 0, TRUE to 0.5)
)

# title and legend
#legend("topleft", 
#       legend=paste(levels(as.factor(Q5.type))), 
#       col = coul , 
#       bty = "n", pch=20 , pt.cex = 2, cex = 1,
#       text.col="black" , horiz = F)



Q5.ceb <- cluster_edge_betweenness(Q5.network) 

dendPlot(Q5.ceb, mode="hclust")

edge.start <- ends(Q5.network, es=E(Q5.network), names=F)[,1]
edge.col <- V(Q5.network)$color[edge.start]

plot(Q5.ceb, Q5.network, axes=F, ,vertex.shape="none",edge.color=edge.col, edge.curved=.1,
     vertex.label.color="black", vertex.label.font=1,vertex.label.dist=0.5,layout=layout_with_fr(Q5.network)) 



```




```{r,echo=FALSE,warning=FALSE,message=FALSE}
edgebundle(Q5.network, padding = 120, fontsize = 10)

```

## Convergence vs. divergence of themes over time

Once again, we binned scores in the the same four groups as above: June, July, Sept/early Oct, and second part of Oct-early Dec.\


```{r,echo=FALSE,warning=FALSE,message=FALSE}

Q5.data.M<-all.data %>% group_by(M)%>%
  summarise(across(balance.of.personal...others.interest_q5:personal.resilience_q5, sum,na.rm=T))


colnames(Q5.data.M)[colnames(Q5.data.M)      # Rename two variable names
                     %in% Q5] <- Q5.q

Q5.data.long.M <- Q5.data.M %>% 
  pivot_longer(-M,names_to = "Theme", values_to = "Count")
Q5.summ.M<-cbind(Q5.data.long.M, Q5.type)
colnames(Q5.summ.M) <- c("Period","Theme","Count","Type")  # change column names
Q5.summ.M <- Q5.summ.M[order(Q5.summ.M$Count), ]  # sort

Q5.data.T<-Q5.summ.M %>% group_by(Period, Type)%>%
    summarise(score=sum(Count))

countByMonth<-all.data %>% group_by(M)%>% 
  summarise(count = n() )

colnames(countByMonth) <- c("Period","N")  # change column names

Q5.data.TT<-full_join(Q5.data.T,countByMonth, by = "Period")
Q5.data.TT$Percent<-(Q5.data.TT$score/Q5.data.TT$N)*100

ggplot(Q5.data.TT, aes(x=Period, y=Percent, group = Type, color=Type, fill=Type)) + 
    geom_line( ) +
    geom_point(shape=21,  size=4) +
   theme_pubclean()+scale_color_futurama()+scale_fill_futurama()

#BY EACH THEME
#Q5.data.T.ALL<-all.data %>% group_by(M)%>%
#   summarise_at(Q5, funs(sum))
#colnames(Q5.data.T.ALL)[colnames(Q5.data.T.ALL)      # Rename variable names
#                  %in% Q5] <- Q5.q
#
#Q5.data.T.ALL.long <-Q5.data.T.ALL %>%
#    gather(domain, frequency, -c(M), convert = TRUE)
#colnames(Q5.data.T.ALL.long) <- c("Period","Theme","Count")  # change column names
#Q5.data.TT.ALL.long<-full_join(Q5.data.T.ALL.long,countByMonth,, by = "Period")
#Q5.data.TT.ALL.long$Percent<-(Q5.data.TT.ALL.long$Count/Q5.data.TT.ALL.long$N)*100
#ggplot(Q5.data.TT.ALL.long, aes(x=Period, y=Percent, group = Theme, color=Theme, fill=Theme)) + 
#    geom_line( ) +
#  geom_point(shape=21,  size=4) +
#  theme_pubclean()

```

Focus on meta-cognition is the top category across the summer and early fall, but becomes second frequent in late fall.\
Focus on well-being enhancement and agency/control are more prevalent among researchers surveyed in late fall. \

# Total wisdom frequencies across questions

When examining categories over time, it becomes apparent that most frequent themes for wisdom either concern interpersonal prosociality (solidarity, social connectedness) or societal level-prosocility (asking for structural changes for a fair and just society), as well as meta-cognitive strategies (perspective-taking, sympathy & compassion, long-term focus, acknowledge unceryainty, self-distancing, critical thinking)

```{r,echo=FALSE,warning=FALSE,message=FALSE}
# STACKED BAR CHART for wisdom
Q2.summ <- Q2.summ[order(Q2.summ$Percent), ]  # sort
Q2.summ$Theme <- factor(Q2.summ$Theme, levels = Q2.summ$Theme)  # to retain the order in plot

Q4.summ <- Q4.summ[order(Q4.summ$Percent), ]  # sort
Q4.summ$Theme <- factor(Q4.summ$Theme, levels = Q4.summ$Theme)  # to retain the order in plot


Q5.summ <- Q5.summ[order(Q5.summ$Percent), ]  # sort
Q5.summ$Theme <- factor(Q5.summ$Theme, levels = Q5.summ$Theme)  # to retain the order in plot

data.245<-full_join(Q4.summ,Q2.summ,by=c("Theme"))
data.245<-data.frame(data.245$Theme,data.245$Percent.x,data.245$Percent.y)
colnames(data.245)<-c("Theme","Percent.Q4","Percent.Q2")
data.245.j<-full_join(data.245,Q5.summ,by=c("Theme"))
data.245.j<-data.frame(data.245.j$Theme,data.245.j$Percent.Q2,data.245.j$Percent.Q4,data.245.j$Percent)
colnames(data.245.j)<-c("Theme","Wisdom\nfor Positive","Wisdom\nagainst Negative","Wisdom Now")
data.245.j[is.na(data.245.j)]<-0


## Reorder fullname based on the the sum of the other columns
data.245.j$Theme <- reorder(data.245.j$Theme, rowSums(data.245.j[-1]))

md <- melt(data.245.j, id=(c("Theme")))
colnames(md)<-c("Theme","Variable","Percent")

g<-ggplot(md, aes(x =Percent , y=Theme,fill = Variable))+geom_bar(stat = "identity")+theme_pubclean()+scale_fill_d3()+theme(legend.position="bottom",legend.title = element_blank())

fig <- ggplotly(tooltip=c("Percent"), height=1020, width = 690)
fig <- fig %>% layout(autosize = T) %>% layout(legend = list(x = 0.6, y = 0.07))

fig


#graph for the infographic
data.245.jS<-data.245.j %>% 
  arrange(desc(rowSums(data.245.j[-1]))) %>% 
  slice(1:11) 


mdS <- melt(data.245.jS, id=(c("Theme")))
colnames(mdS)<-c("Theme","Variable","Percent")

gS<-ggplot(mdS, aes(x =Percent , y=Theme,fill = Variable), width = 0.5)+
  geom_bar(stat = "identity")+ 
  scale_color_tableau() +theme_minimal() +
  theme(legend.position="bottom")+
  scale_fill_manual(values=c("#F28E2B", "#4E79A7",  "#59A14F"))+
  ylab("")+
  xlab("%") +
  theme(legend.title = element_blank())




```


# Does dialecticism and knowledge of wisdom scholarship impact type of wisdom experts recommended?

## Dialectical thinking

So far, we just examined relative frequencies for certain wisdom-related themes, separately for each question. We did not account for interdependence between responses by the same person nor did we look at actual likelihood of mentioning one vs. another category (e.g., moral vs. others or metacog vs. others). To do the latter, I restructure the dataset to a long format and perform generalized (binomial) linear mixed model with dichotomous responses (1/0) nested in participants and type of wisdom theme (moral vs. metacog vs. other), cognitive style (dialectical vs. non-dialectical) and their interaction as predictors. If cognitive style of experts qualifies type of category mentioned, we should expect a significant interaction.\
\
It turns out, it does not matter, with no significant differences for any of the questions or in total.\

```{r,warning=FALSE,message=FALSE}
#restructure to long format

wisdom.total<-c("social.connectedness_q2", "social.support_q2",  "evidence.based.decision.making_q2","improved.work.life.balance_q2","acknowledge.of.uncertainty.flexibility_q2","balance.of.personal...others.interest_q2","perspective.taking_q2",
        "critical.thinking_q2","bipartisanship.and.international.cooperation_q2","solidarity_q2",                        
        "improve.communication_q2", "intellectual.humility_q2","living.in.the.moment_q2",
        "learning.from.pandemics_q2","political.engagement...structural.change_q2",
        "self.reflection.on.what.s.important_q2","shared.humanity_q2","sympathy...compassion_q2", "self.distancing_q2","embrace.new.tech_q2","personal.resilience_q2",
        "social.awareness_q4" , "balance.of.personal...others.interest_q4" ,  "solidarity_q4",                    "context.sensitivity_q4",                          "bipartisanship.and.international.cooperation_q4", "critical.thinking_q4",                            "evidence.based.decision.making_q4",               "gratitude_q4",                                   
        "optimism...positivity_q4",                              "improve.communication_q4",                        "long.term.orientation_q4",    "social.support_q4"  ,                  "social.connectedness_q4",               "acknowledge.uncertainty.flexibility_q4",         
        "living.in.the.moment_q4",                         "patience_q4",                                     "perspective.taking_q4",                           "self.reflection.on.what.s.important_q4",          "shared.humanity_q4",                             
        "strive.for.socio.economic.equality_q4",           "sympathy...compassion_q4",                        "political.engagement...structural.change_q4",     "self.distancing_q4",
        "balance.of.personal...others.interest_q5",       
 "clear.governmental.communication_q5",             "solidarity_q5",                                   "bipartisanship.and.international.cooperation_q5",
 "critical.thinking_q5",                            "embrace.new.tech_q5",                             "agency...control_q5",                            
 "follow.rules_q5",                                 "optimism...positivity_q5",                        "learning.from.the.pandemics_q5",                  "improve.communication_q5",
 "prosocial.behavior_q5",                           "long.term.orientation_q5",         "acknowledge.uncertainty.flexibility_q5",          "living.in.the.moment_q5",
 "patience_q5",                                     "perspective.taking_q5",                           "self.reflection.on.what.is.important_q5",         "social.connectedness_q5",                        
 "self.distancing_q5",                              "strive.for.socio.economic.equality_q5",           "sympathy...compassion_q5",                        "take.science.seriously_q5",                       "personal.resilience_q5", "Dialectic_Final","FamiliarWisdom","Name")

wisdom.long<-all.data[wisdom.total] %>% gather(key = "theme", value = "value", -Dialectic_Final,-FamiliarWisdom,-Name)

morals.Q2<-enframe(morals.Q2)
morals.Q2$Q<-"Q2"
morals.Q4<-enframe(moralsQ4)
morals.Q4$Q<-"Q4"
morals.Q5<-enframe(moralsQ5)
morals.Q5$Q<-"Q5"

morals<-rbind(morals.Q2,morals.Q4,morals.Q5)
morals$theme<-morals$value
morals$type<-"moral"

metacog.Q2<-enframe(metacog.Q2)
metacog.Q2$Q<-"Q2"
metacog.Q4<-enframe(metacogQ4)
metacog.Q4$Q<-"Q4"
metacog.Q5<-enframe(metacogQ5)
metacog.Q5$Q<-"Q5"
metacog<-rbind(metacog.Q2,metacog.Q4,metacog.Q5)
metacog$theme<-metacog$value
metacog$type<-"metacog"

others.Q2<-enframe(others.Q2)
others.Q2$Q<-"Q2"
others.Q4<-enframe(othersQ4)
others.Q4$Q<-"Q4"
others.Q5<-enframe(othersQ5)
others.Q5$Q<-"Q5"
others<-rbind(others.Q2,others.Q4,others.Q5)
others$theme<-others$value
others$type<-"others"

scores<-select(rbind(morals,metacog,others),-c("name","value"))

wisdom.long.total<-wisdom.long %>% left_join(scores,by="theme")

#Run analyses

## Does it matter if the person is dialectical?

dialect.test<-glmer(value~Dialectic_Final*type*Q+(1|Name),data=wisdom.long.total,family=binomial)
car::Anova(dialect.test,type="III")   #no significant effect and no interaction 

### test by question: positive

dialect.test.q2<-glmer(value~Dialectic_Final*type+(1|Name),data=wisdom.long.total %>% filter(Q=="Q2"),family=binomial)
car::Anova(dialect.test.q2,type="III")   #no significant effect and no interaction 

### test by question: negative

dialect.test.q4<-glmer(value~Dialectic_Final*type+(1|Name),data=wisdom.long.total %>% filter(Q=="Q4"),family=binomial)
car::Anova(dialect.test.q4,type="III")   #no significant effect and no interaction 

### test by question: now

dialect.test.q5<-glmer(value~Dialectic_Final*type+(1|Name),data=wisdom.long.total %>% filter(Q=="Q5"),family=binomial)
car::Anova(dialect.test.q5,type="III")   #no significant effect and no interaction 




all.data$morals.QALL<-rowSums(all.data[c("morals.Q2","morals.Q4","morals.Q5")])
all.data$metacog.QALL<-rowSums(all.data[c("metacog.Q2","metacog.Q4","metacog.Q5")])
all.data$others.QALL<-rowSums(all.data[c("others.Q2","others.Q4","others.Q5")])

all.data$CWM.QALL<-rowSums(all.data[c("morals.QALL","metacog.QALL")])

all.dia.data<-all.data %>% pivot_longer(cols=morals.QALL:others.QALL, names_to="type",values_to="count")

all.dia.result<-glmer(count ~ type*Dialecticism + (1 | Name), data = all.dia.data, family = poisson)
Anova(all.dia.result,type="III") #no significant differences
sjPlot::plot_model(all.dia.result,type="int")

all.dia.result.emt<-emmeans::emmeans(all.dia.result,"type",by="Dialecticism")
pairs(all.dia.result.emt)

pairs(emmeans::emmeans(all.dia.result,"Dialecticism",by="type")) #no sig difference in total

#contrast of cWM (morals and meta-cog) vs. the rest

ALL.dia.data.C<-all.data %>% pivot_longer(cols=c("others.QALL","CWM.QALL"), names_to="type",values_to="count")

QALL.dia.result.C<-glmer(count ~ type*Dialecticism + (1 | Name), data = ALL.dia.data.C, family = poisson)
Anova(QALL.dia.result.C,type="III") #no significant differences
sjPlot::plot_model(QALL.dia.result.C,type="int")

QALL.dia.result.C.emt<-emmeans::emmeans(QALL.dia.result.C,"type",by="Dialecticism")
pairs(QALL.dia.result.C.emt)
pairs(emmeans::emmeans(QALL.dia.result.C.emt,"Dialecticism",by="type")) #no sig difference in total




```





## Does familiarity with wisdom research influence types of categories experts mention in their recommendations?

To address this question, I perform generalized (binomial) linear mixed model with dichotomous responses (1/0) nested in participants and type of wisdom theme (moral vs. metacog vs. other), familiarity with wisdom (yes vs. no) and their interaction as predictors. If familiarity with wisdom qualifies type of category mentioned, we should expect a significant interaction.\
\
It turns out, it does not matter for positive recommendations and for recommendations now (as well as in total), though we did observe a significant interaction for wisdom against negative consequences: Experts familiar with wisdom scholarship were significantly more likely to mention meta-cognition than other categories, whereas no such differences among those folks not familiar with wisdom scholarship.\

```{r,echo=FALSE,warning=FALSE,message=FALSE}

## Does it matter if the person is familiar with wisdom research (supplementary question)?

wis.test<-glmer(value~FamiliarWisdom*type*Q+(1|Name),data=wisdom.long.total,family=binomial)
car::Anova(wis.test,type="III")   #no significant effect and no interaction 


QALL.famil.result<-glmer(count ~ type*FamiliarWisdom + (1 | Name), data = all.dia.data, family = poisson)
Anova(QALL.famil.result,type="III") #no significant differences
sjPlot::plot_model(QALL.famil.result,type="int")

QALL.famil.result.emt<-emmeans::emmeans(QALL.famil.result,"type",by="FamiliarWisdom")
pairs(QALL.famil.result.emt)

pairs(emmeans::emmeans(QALL.famil.result,"FamiliarWisdom",by="type"))


### test by question: positive

wisdom.test.q2<-glmer(value~FamiliarWisdom*type+(1|Name),data=wisdom.long.total %>% filter(Q=="Q2"),family=binomial)
car::Anova(wisdom.test.q2,type="III")   #no significant effect and no interaction 

### test by question: negative

wisdom.test.q4<-glmer(value~FamiliarWisdom*type+(1|Name),data=wisdom.long.total %>% filter(Q=="Q4"),family=binomial)
car::Anova(wisdom.test.q4,type="III")   # significant effect and a sig interaction 
sjPlot::plot_model(wisdom.test.q4, type="int") #familiar with wisdom are significantly more likely to mention meta-cognition than moral aspirations and (esp. others), whereas no such differences among those folks not familiar with wisdom scholarship

### test by question: now

wisdom.test.q5<-glmer(value~FamiliarWisdom*type+(1|Name),data=wisdom.long.total %>% filter(Q=="Q5"),family=binomial)
car::Anova(wisdom.test.q5,type="III")   #no significant effect and no interaction 

```

# Relationship of Forecasts & Given Advice

A key question concerns type of advice scientists recommend for different (positive and negative forecasts) for positive and negative changes after the pandemic.\
\
To address this question, we can examine **dependencies** between themes mentioned for a particular type of outcome, and subsequent mentioning of themes needed for this outcome.\


## Positive Consequences

Given that some themes are mentioned *both* as outcomes of the pandemic (i.e., forecasts) and the advice/wisdom needed to sustain positive change (e.g., critical thinking, live in the moment, political/structural change), it would be trivial to see relevant relationships. However, it would be more interesting to see relationships *across* themes from forecast (Q1) and corresponding advice (Q2).\


### Heatmap

One way to address this question is to examine the heatmap, with forecast themes on the vertical axis and the advice themes on the horizontal axis.\
\
As one can see below, some of the results are indeed quite trivial. Beyond same themes mentioned for forecasts and advice about wisdom needed, another trivial observation is that greater care for elderly depends on *willingness to learn from the pandemic*. Nevertheless, some further observations are noteworthy. I focus on r > .3 as a cut off:\

* greater science interest in the future depends on *critical thinking*;\
* greater social connectedness and optimism/positivity depend on *improved communication*;\
* greater health & wellbeing depends on *improved work-life balance* , *focus on living in the moment* & *ability to compromise/balance diverse interests*;\
* greater resilience depends on *sympathy & compassion*;\
* greater resilience depends on *sympathy & compassion*;\
* learning from the pandemic in the future depends on *political cooperation*;\
* greater solidarity depends on *awareness of shared humanity*;\
* greater transition to new technologies and reconsideration of habits depend on *personal resilience*;\
* **strongest relationship**: greater resilience, embacing new tech, and reconsideration of habits depend on *acknowledgement of uncertainty*.


```{r,echo=FALSE,warning=FALSE,message=FALSE, out.height="300%"}

##Heatmap

#create a subset of the matrix with Q1 in rows and Q2 in columns

Q12<-c("shared.humanity_q1","critical.thinking_q1","personal.resilience_q1","gratitude_q1",
       "nature_q1","improved.care.for.elders_q1","embrace.new.tech_q1","prosocial.behavior_q1",
       "social.connectedness_q1","improved.work.life.balance_q1","increased.interest.in.science_q1",
       "learning.from.pandemics_q1","political.engagement...structural.change_q1",
       "optimism...positivity_q1","reconsidering.habits_q1",
       "living.in.the.moment_q1","health...well.being_q1","solidarity_q1","bipartisanship.and.international.cooperation_q1","sympathy...compassion_q1",
       "social.connectedness_q2", "social.support_q2",  "evidence.based.decision.making_q2","improved.work.life.balance_q2","acknowledge.of.uncertainty.flexibility_q2","balance.of.personal...others.interest_q2","perspective.taking_q2",
       "critical.thinking_q2","bipartisanship.and.international.cooperation_q2","solidarity_q2",                        
       "improve.communication_q2", "intellectual.humility_q2","living.in.the.moment_q2",
       "learning.from.pandemics_q2","political.engagement...structural.change_q2",
       "self.reflection.on.what.s.important_q2","shared.humanity_q2","sympathy...compassion_q2", "self.distancing_q2","embrace.new.tech_q2","personal.resilience_q2") 

Q12.cor.data<-partial.r(all.data,Q12,cs(tokens.Q1,tokens.Q2),method="spearman") #correct for tokens (length) in each interview
write.csv(Q12.cor.data,"corr12.csv") #R is weird with partial.r function so I jsut save and open the file again, scripping it to one layer
Q12.cor.data<-read.csv("corr12.csv")

Q12.cor.arb<-Q12.cor.data[1:length(Q1),c("X",Q2)]
Q12.cor.arb <- data.frame(Q12.cor.arb[,-1], row.names=paste("Q1",Q1.q))
colnames(Q12.cor.arb) <-paste("Q2",Q2.q)

heatmaply_cor(round(Q12.cor.arb,digits=2), #round to 2 decimal points
              #dendrogram = "row",
              xlab = "", ylab = "", plot_method="ggplot",fontsize_row = 8,
  fontsize_col = 8,
              main = "",
              distfun="spearman",
              #scale = "column",
              grid_color = "white",
            #  grid_width = 0.00001,
              titleX = FALSE,
              hide_colorbar = T,
             # branches_lwd = 0.1,
              label_names = c("Q1", "Q2:", "Value"),
              #fontsize_row = 5, fontsize_col = 5,
              labCol = colnames(Q12.cor.arb),dendrogram = F, showticklabels = T, margins = c(1,1),
              labRow = rownames(Q12.cor.arb),
              heatmap_layers = theme(axis.line=element_blank()))%>% layout(margin = list(l = 130, b = 40))
```

### Q1-Q2 Edge Binding

A different, and perhaps more intuitive way to examine the relationship between themes mentioned for forecasts and advice is to examine connections between "edges" formed in the network of QUestions concerning forecasts and the advice. Here, full information from the network is used to optimally visualize connections between different themes in the network.\
\
Below, you see an edge-binding diagram visualizing the relationships. On the right, you see themes for forecasts (Q1) and on the left you see type of advice recommended (Q2). When you navigate the cursor on each theme, the graph will show you what other themes it is mostly connected to, both within the same question and across questions. As for the other network visualizations, I used r < .17 as a cut-off here.\
\
For instance, if you navigate to *Q1_Gratitude*, you will see that it is related not only to Q1_living in a moment and Q1_new tech, but also Q2 focus on what's important, Q2 living in the moment, Q2 critical thinking, and Q2 balance of diverse interests. \
\
By inspecting all Q1 themes, we can see that themes followed by greatest number of advice-themes are:\

* Gratitude;\
* Greater appreciation of Nature;\
* Resilience;\
* Solidarity.\
\
Conversely, by inspecting all Q2 themes, we can see that themes with greatest number of implications for forecasts are:\

* Acknowledgment of uncertainty;\
* Solidarity;\
* Self-distancing.\


```{r,echo=FALSE,warning=FALSE,message=FALSE}

#edge-bundling
Q12.cor<-partial.r(all.data,Q12,cs(tokens.Q1,tokens.Q2),method="spearman")


Q12.q <- c("Q1_Shared Humanity","Q1_Critical Thinking","Q1_Resilience","Q1_Gratitude","Q1_Nature","Q1_Care for Elders","Q1_Embrace New Tech","Q1_Prosocial Behavior",
           "Q1_Social Connectedness","Q1_Work-Life Balance","Q1_Science Interest","Q1_Learning from Pandemics","Q1_Polit Change",
           "Q1_Optimism/Positivity","Q1_Reconsider Habits","Q1_Live\nin the Moment","Q1_Health/Wellbeing","Q1_Solidarity","Q1_Polit\nCooperation","Q1_Sympathy&Compassion",
           "Q2_Social Connectedness", "Q2_Social Support", "Q2_Evidence-based Judgement", "Q2_Work-Life Balance", "Q2_Uncertainty","Q2_Balance Interests","Q2_Perspective-taking",
           "Q2_Critical Thinking","Q2_Polit\nCooperation",
           "Q2_Solidarity", "Q2_Improved Communication", "Q2_Intellectual Humility","Q2_Live in the Moment",
           "Q2_Learning from Pandemics","Q2_Polit Change","Q2_What's Important?","Q2_Shared Humanity","Q2_Sympathy&Compassion", "Q2_Self-distancing","Q2_Embrace New Tech","Q2_Personal Resilience") 

row.names(Q12.cor) <-Q12.q
colnames(Q12.cor) <-Q12.q

# Keep only meaningful correlations (<.19)
Q12.cor[Q12.cor<.17] <- 0


# Make an Igraph object from this matrix:
Q12.network <- graph_from_adjacency_matrix( Q12.cor, weighted=T, mode="undirected", diag=F)

edgebundle(Q12.network, padding = 120, fontsize = 10) #revise so that Q1 / Q2 appears first.


```

## Negative Consequences

We can do the same analysis for Questions 3 (negative forecasts) and 4 (wisdom needed to prevent these forecasts).\

### Heatmap

Again, I will start with a heatmap, with forecast themes on the vertical axis and the advice themes on the horizontal axis.\
\
Given the negative nature of the forecasts, we would not expect as many trivial results (except for things like: who do you need to reduce socio-econ or edu equality? Fight against socio-econ inequality! Or: "how do we prevent dispair? Be optimistic/positive :)). Once again, I focus on r > .3 as a cut off:\

* combating estrangement/alienation, loneliness, decline in wellbeing was linked to the notion of *social connectedness*;\
* combating economic hardships and rise in despair was linked to *social support*;\
* combating rising social inequality was linked to *solidarity*;\
* **strongest relationships**: combating problems in intimate relations, educational inequality, and child development issues were linked to *ability to balance/reach a compromise across diverse interests*;\
* combating decline in autobiographic memory, estrangement/alienation, and well-being were linked to *contextualizing one's life experiences*;\
* combating irrationality depends on *living in a moment/mindfulness*,*gratitude*, and *self-distancing*;\
* combating low trust in science was linked to *self-distancing*;\
* combating political conflict was linked to *appreciation of the concept of shared humanity*;\



```{r,echo=FALSE,warning=FALSE,message=FALSE}

##Heatmap

#create a subset of the matrix with Q3 in rows and Q4 in columns

Q34<-c("estrangement...alienation_q3","career.disruptions_q3", "decline.in.quality.of.education_q3", "decreased.trust.in.science_q3","decreased.well.being_q3"   ,                     
        "economic.hardship_q3" ,"educational.inequality_q3", "erosion.of.democratic.institutions_q3", "exacerbated.social.inequality_q3","mistrust_q3",                            
        "prejudice...racism_q3", "despair_q3", "prioritize.self.over.others_q3","irrationality_q3","loneliness_q3"  ,                                
        "misperception.about.the.world_q3" ,"political.conflict_q3" , "authoritarianism_q3", "stunted.child.development_q3","pessimism_q3",                                   
        "intimate.relations_q3" , "autobiographical.memory_q3","social.awareness_q4" , "balance.of.personal...others.interest_q4" ,  "solidarity_q4",                    "context.sensitivity_q4",                          "bipartisanship.and.international.cooperation_q4", "critical.thinking_q4",                            "evidence.based.decision.making_q4",               "gratitude_q4",                                   
        "optimism...positivity_q4",                              "improve.communication_q4",                        "long.term.orientation_q4",    "social.support_q4"  ,                  "social.connectedness_q4",               "acknowledge.uncertainty.flexibility_q4",         
        "living.in.the.moment_q4",                         "patience_q4",                                     "perspective.taking_q4",                           "self.reflection.on.what.s.important_q4",          "shared.humanity_q4",                             
        "strive.for.socio.economic.equality_q4",           "sympathy...compassion_q4",                        "political.engagement...structural.change_q4",     "self.distancing_q4")

Q34.cor.data<-partial.r(all.data,Q34,cs(tokens.Q3,tokens.Q4),method="spearman") #correct for tokens (length) in each interview
write.csv(Q34.cor.data,"corr34.csv") #R is weird with partial.r function so I jsut save and open the file again, scripping it to one layer
Q34.cor.data<-read.csv("corr34.csv")

Q34.cor.arb<-Q34.cor.data[1:length(Q3),c("X",Q4)]
Q34.cor.arb <- data.frame(Q34.cor.arb[,-1], row.names=paste("Q3",Q3.q))
colnames(Q34.cor.arb) <-paste("Q4",Q4.q)

heatmaply_cor(round(Q34.cor.arb,digits=2), #round to 2 decimal points
              #dendrogram = "row",
              xlab = "", ylab = "", plot_method="ggplot",fontsize_row = 8,
  fontsize_col = 8,
              main = "",
              distfun="spearman",
              #scale = "column",
              grid_color = "white",
            #  grid_width = 0.00001,
              titleX = FALSE,
              hide_colorbar = T,
             # branches_lwd = 0.1,
              label_names = c("Q3", "Q4:", "Value"),dendrogram = F, 
              #fontsize_row = 5, fontsize_col = 5,
              labCol = colnames(Q34.cor.arb),margins = c(1,1),
              labRow = rownames(Q34.cor.arb),
              heatmap_layers = theme(axis.line=element_blank()))%>% layout(margin = list(l = 150, b = 70))
```

### Q3-Q4 Edge Binding

Once again, below you see an edge-binding diagram  - a different way to visualize the relationships. On the right, you see themes for *negative* forecasts (Q3) and on the left you see type of advice recommended (Q4). When you navigate the cursor on each theme, the graph will show you what other themes it is mostly connected to, both within the same question and across questions. As for the other network visualizations, I used r < .17 as a cut-off here.\
\
For instance, if you navigate to *Q1_Gratitude*, you will see that it is related not only to Q1_living in a moment and Q1_new tech, but also Q2 focus on what's important, Q2 living in the moment, Q2 critical thinking, and Q2 balance of diverse interests. \
\
By inspecting all Q3 themes, we can see that themes followed by greatest number of advice-themes are:\

* Estrangement/alienation;\
* Despair;\
* Self-centeredness;\
* Irrationality;
* Loneliness;\
* Pessimism;\


Conversely, by inspecting all Q4 themes, we can see that themes with greatest number of implications for forecasts are:\

* Self-distancing;\
* Political/structural change;\
* Perspective-taking;\
* Patience;\
* Gratitude.\


```{r,echo=FALSE,warning=FALSE,message=FALSE}

#edge-bundling
Q34.cor<-partial.r(all.data,Q34,cs(tokens.Q3,tokens.Q4),method="spearman")


Q34.q <- c("Q3_Estrangement","Q3_CareerDisruptions","Q3_Low QualEducation","Q3_LowTrust\nin Science","Q3_LowWellbeing"   ,                     
          "Q3_EconomicHardship" ,"Q3_EducationalInequality","Q3_ErosionofDemoc","Q3_SocialInequality","Q3_Mistrust",                            
          "Q3_PrejudiceRacism","Q3_Despair","Q3_Self-Centeredness","Q3_Irrationality","Q3_Loneliness"  ,                                
          "Q3_MispercWorld" ,"Q3_PolitConflict" ,"Q3_Authoritarianism","Q3_StuntedDev","Q3_Pessimism",                                   
          "Q3_IntimateRelations" ,"Q3_AutobioMemory","Q4_SocialAwareness" ,"Q4_BalanceInterests" ,   "Q4_Solidarity"    ,                    
          "Q4_ContextSensitivity",  "Q4_PolCoop","Q4_CriticalThinking",  
          "Q4_Evidence-bJudgement","Q4_Gratitude","Optimism/Positivity","ImprovedComm","Q4_Long-termOrien",   "Q4_SocialSupport","Q4_SocialConnectedness",
          "Q4_Uncertainty",  "Q4_LivetheMoment","Q4_Patience", "Q4_Persp-taking","Q4_What's Important?","Q4_SharedHumanity",                             
          "Q4_Socio-econEquality",           "Q4_Sympathy&Compassion", "Q4_PolitChange",     "Q4_Self-distancing") 

row.names(Q34.cor) <-Q34.q
colnames(Q34.cor) <-Q34.q

# Keep only meaningful correlations (<.19)
Q34.cor[Q34.cor<.17] <- 0


# Make an Igraph object from this matrix:
Q34.network <- graph_from_adjacency_matrix( Q34.cor, weighted=T, mode="undirected", diag=F)

edgebundle(Q34.network, padding = 130, fontsize = 9) #revise so that Q1 / Q2 appears first.


```